### 2023-08-09 Paper/paper.Rmd
Started writing on the paper, minor changes in baseline regressions

---
title: "The Gender Pay Gap in the General Social Survey"
author: "Alex Amaguaya, Lea Bergmann"
cols_authors: 3
subtitle: "Statistical Learning"
deadline: "tomorrow"
type: "Term Paper"
date: "today"
supervisors: "Prof. Dr. Thomas Deckers, Karolina Gliszczynska"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: references.bib
classoption: a4paper
---

<!-- % Template Version 1.1 -->

<!-- below function does some formatting for images; leave this untouched unless you know better :-) -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magick)
library(tidyverse)
library(stargazer)
crop <- function(im, left = 0, top = 0, right = 0, bottom = 0) {
  d <- dim(im[[1]]); w <- d[2]; h <- d[3]
  image_crop(im, glue::glue("{w-left-right}x{h-top-bottom}+{left}+{top}"))
}
```

# Introduction

In this paper we predict individuals' wages using data from the General Social Survey (GSS, 1974-2018). The variable of interest is `realrinc` which varies between 227\$ and 480,145\$. Hence, we can use regression techniques to predict individuals' wages. Additionally, we can consider whether `realrinc` varies by gender and if so, how much.

The data from the GSS allow us to consider various control variables such as gender, age, education, occupation, number of children and marital status.

To predict `realrinc` we use several regression techniques:

-   Linear Regression

-   Logistic Regression

-   Support Vector Machines

-   Gradient Boosting

-   Random Forest

In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%), a validation (15%) and a test data set (15%). We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RSME) using the validation data.

[Write something about why to use RSME]

[Write something about cross validation?]

# Data Preparation

Before conducting the regression analysis we prepared the data. The provided dataset has 54,842 observations and for 38.55% of the observations the outcome variable `realrinc` is missing. As we cannot predict the wage for those observations we removed them. This left us with a dataset containing 33,702 observations. Additionally, we do not know the occupation for some observations (`occ10 = 9997` ) and removed those. This diminishes the dataset to 33,244 observations.

In this dataset still there were missing some values for some observations, e.g., number of children or occupation. To impute these values we used the `mice` function. We imputed age, occupational prestige and number of children with predictive mean matching. For the factor variables education and marital status we used polytomous logisitic regression. We receive six imputation datasets and use mean estimation for age, occupational prestige and number of children.

With this dataset we create dummy variables for the factor variables. Additionally, we create interaction variables.

[Something about ANOVA test for interaction variables].

In a final step we calculate the correlation matrix using Spearman's rank correlation coefficient. We remove all variables that have a higher than 90% correlation with each other. Our final dataset then contains 33,233 of 347 variables.

# Model Training and Tuning

We used five different regression techniques to predict the individuals' wages using the final dataset. We fit every model on our training data using the `train` function from the `caret` package. To choose the best parameter combination we use cross-validation within all train functions.

## Linear and Logistic Regression

Our baseline linear model uses an OLS regression including all interaction variables. This results in an RMSE of 20,075.61.

``` R
mod_full <- train(realrinc ~  . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
                  data = train, 
                  method = "lm",  
                  trControl = trainControl(method = "cv"))
```

Given some extreme outliers in the income distribution of our data (which is typical for wage data) we follow the economic literature ([https://onlinelibrary.wiley.com/toc/14680084/2008/70/s1](#0)) [correct citation missing] and fit a logistic model to our data, using the log of `realrinc` as our dependent variable. This results in a higher RMSE than the linear model (35,844.61).

``` R
mod_loginc <- train(log_realrinc ~ . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
                    data = train, 
                    method = "lm",  
                    trControl = trainControl(method = "cv")) 
```

In both models the coefficient of female remains statistically insignificant at conventional levels. This suggests that gender is not a driving factor for wages within our data.

## Support Vector Machines

### 2023-08-09 code/1_3_data_preparation.R

@@ -265,8 +265,10 @@ full_data_final <- cbind(ds.w.imputed.dummies,
                          mx.result.tibble) %>% as_tibble()
 
 # fast estimation of correlation matrix
-cor.filter <- select_if(ds.w.imputed.dummies, is.numeric)
-cor(cor.filter, method = "spearman")
+cor.filter <- select_if(ds, is.numeric)
+matrix <- cor(cor.filter, method = "spearman")
+highCorFeatures <- findCorrelation(matrix, cutoff = 0.9, exact = TRUE)
+ds <- ds[,-highCorFeatures]
 
 # save datasets
 write.csv(full_data_final, file = "data/full_data_w_dummies_interaction.csv")

### 2023-08-09 code/2_baseline_regressions.R

@@ -26,7 +26,7 @@ test <- remaining[-valid_test_index, ]
 
 
 #OLS regressions
-mod_full <- train(realrinc ~ . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
+mod_full <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
                   data = train, 
                   method = "lm",  
                   trControl = trainControl(method = "cv"))
@@ -51,7 +51,7 @@ cat("Mean Squared Error (MSE):", mse, "\n") #605727520
 rmse <- caret::RMSE(pred = predict(mod_wo_interactions, validation), obs = validation$realrinc)
 cat("Root Mean Squared Error (RMSE):", rmse, "\n") #24611.53
 
-mod_loginc <- train(log_realrinc ~ . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
+mod_loginc <- train(log_realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
                     data = train, 
                     method = "lm",  
                     trControl = trainControl(method = "cv"))

### 2023-08-09 code/3_SupportVector.R

@@ -28,7 +28,7 @@ test <- remaining[-valid_test_index, ]
 lin_kern <- train(realrinc ~ age + prestg10 + childs + armed_forces + business_finance + construction_extraction + farming_fishing_and_forestry + 
                     installation_maintenance_and_repair + office_and_administrative_support + production + professional + sales + service + 
                     + full_time + housekeeper + part_time + retired + school + temporarily_not_working + unemployed_laid_off + bachelor + 
-                    graduate + highschool + juniorcollege +  + divorced + married + nevermarried + separated  + female, data = train, 
+                    graduate + highschool + juniorcollege +  divorced + married + nevermarried + separated  + female, data = train, 
       method = 'svmLinear', preProcess = c("center","scale"),
       trControl = trainControl(method = "cv"), tuneGrid = data.frame(C = c(0.01,0.1,1,10,100)) )
 rmse <- caret::RMSE(pred = predict(lin_kern, validation), obs = validation$realrinc)

### 2023-08-10 code/4_LassoRidgeElasticNet.R
Lasso and ridge regressions

@@ -1,0 +1,45 @@
+# set wd
+setwd("/home/user/Downloads/info_tud/statistical_learning_UDE/gender_pay_gap_project")
+setwd("C:/Users/lbergmann/OneDrive - RWI–Leibniz-Institut für Wirtschaftsforschung e.V/Dokumente/Promotion/StatisticalLearning/gender_pay_gap_project_1")
+
+# packages
+library(dplyr)
+library(GGally)
+library(caret)
+library(ISLR)
+library(tidyverse)
+library(caret)
+library(glmnet)
+
+
+
+#load data 
+ds <- read.csv("data/full_data_w_dummies_interaction.csv")
+
+#split in train, validation and test data
+set.seed(123)
+
+trainindex <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
+train <- ds[trainindex, ]
+remaining <- ds[-trainindex, ]
+
+valid_test_index <- createDataPartition(y = remaining$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
+validation <- remaining[valid_test_index, ]
+test <- remaining[-valid_test_index, ]
+
+
+
+#Lasso
+lasso <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train, 
+                method = "glmnet", trControl = trainControl(method = "cv"), 
+                tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
+caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #19987.04
+
+
+#Ridge
+ridge <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train,
+  method = "glmnet", trControl = trainControl(method = "cv"),
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #20039.76
+caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
+
+#Elastic Net

### 2023-08-11 code/1_3_data_preparation.R
Improved baseline regressions

@@ -270,5 +270,8 @@ matrix <- cor(cor.filter, method = "spearman")
 highCorFeatures <- findCorrelation(matrix, cutoff = 0.9, exact = TRUE)
 ds <- ds[,-highCorFeatures]
 
+
+full_data_final <- subset(ds, select = -c(row_w_na, educcat, maritalcat, occrecode, wrkstat, gender, X, X.1) )
+
 # save datasets
 write.csv(full_data_final, file = "data/full_data_w_dummies_interaction.csv")

### 2023-08-11 code/2_baseline_regressions.R

@@ -12,6 +12,7 @@ library(caret)
 
 #load data
 ds <- read.csv("data/full_data_w_dummies_interaction.csv")
+ds <- subset(ds, select = -c(X, X.1) )
 
 #split in train, validation and test data
 set.seed(123)
@@ -26,55 +27,121 @@ test <- remaining[-valid_test_index, ]
 
 
 #OLS regressions
-mod_full <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
+mod_full <- train(realrinc ~ female + . - log_realrinc - transportation - unemployed_laid_off - lessthan_high_school - widowed - whiteCollar - age_sqr,
                   data = train, 
                   method = "lm",  
                   trControl = trainControl(method = "cv"))
 summary(mod_full)
-pred <-predict(mod_full, newdata = validation)
-mse <- ModelMetrics::mse(validation$realrinc, pred)
-cat("Mean Squared Error (MSE):", mse, "\n") #403030161
 rmse <- caret::RMSE(pred = predict(mod_full, validation), obs = validation$realrinc)
-cat("Root Mean Squared Error (RMSE):", rmse, "\n") #20075.61 
+cat("Root Mean Squared Error (RMSE):", rmse, "\n") #24429.79
+###I receive warnings due to multicollinearity
 
-mod_wo_interactions <-train(realrinc ~ age + prestg10 + childs + armed_forces + business_finance + construction_extraction + farming_fishing_and_forestry + 
+mod_wo_interactions <-train(realrinc ~ female + age + prestg10 + childs + armed_forces + business_finance + construction_extraction + farming_fishing_and_forestry + 
                             installation_maintenance_and_repair + office_and_administrative_support + production + professional + sales + service + 
                              + full_time + housekeeper + part_time + retired + school + temporarily_not_working + unemployed_laid_off + bachelor + 
-                            graduate + highschool + juniorcollege +  + divorced + married + nevermarried + separated  + female,
+                            graduate + highschool + juniorcollege +  + divorced + married + nevermarried + separated,
                             data = train, 
                             method = "lm",  
                             trControl = trainControl(method = "cv"))
 summary(mod_wo_interactions)
-pred <-predict(mod_wo_interactions, newdata = validation)
-mse <- ModelMetrics::mse(validation$realrinc, pred)
-cat("Mean Squared Error (MSE):", mse, "\n") #605727520 
[…]

### 2023-08-11 code/4_LassoRidgeElasticNet.R

@@ -43,3 +43,8 @@ ridge <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -mar
 caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
 
 #Elastic Net
+elasticnet <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train,
+                    method = "glmnet", trControl = trainControl(method = "cv"),
+                    tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
+                    lambda = seq(from=0, to=0.15, by = 0.001)))
+caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #20045.16

### 2023-08-11 code/4_LassoRidgeElasticNet.R
Adapted Lasso etc.

@@ -15,6 +15,7 @@ library(glmnet)
 
 #load data 
 ds <- read.csv("data/full_data_w_dummies_interaction.csv")
+ds <- subset(ds, select = -c(X, X.1) )
 
 #split in train, validation and test data
 set.seed(123)
@@ -28,23 +29,25 @@ validation <- remaining[valid_test_index, ]
 test <- remaining[-valid_test_index, ]
 
 
-
 #Lasso
-lasso <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train, 
+lasso <- train(realrinc ~ female + . - log_realrinc, data = train, 
                 method = "glmnet", trControl = trainControl(method = "cv"), 
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
-caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #19987.04
+caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #24283.21
 
 
 #Ridge
-ridge <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train,
+ridge <- train(realrinc ~ female + . - log_realrinc, data = train,
   method = "glmnet", trControl = trainControl(method = "cv"),
-  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #20039.76
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #24325.68
 caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
 
 #Elastic Net
-elasticnet <- train(realrinc ~ female + . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr, data = train,
+elasticnet <- train(realrinc ~ female + . - log_realrinc, data = train,
                     method = "glmnet", trControl = trainControl(method = "cv"),
                     tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
                     lambda = seq(from=0, to=0.15, by = 0.001)))
-caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #20045.16
+caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #24333.99
+
+
+write.csv(ds, file = "data/final_data.csv")

### 2023-08-11 code/1_3_data_preparation.R
adapted code regarding x variables

@@ -271,7 +271,5 @@ highCorFeatures <- findCorrelation(matrix, cutoff = 0.9, exact = TRUE)
 ds <- ds[,-highCorFeatures]
 
 
-full_data_final <- subset(ds, select = -c(row_w_na, educcat, maritalcat, occrecode, wrkstat, gender, X, X.1) )
-
 # save datasets
 write.csv(full_data_final, file = "data/full_data_w_dummies_interaction.csv")

### 2023-08-11 code/2_baseline_regressions.R

@@ -9,33 +9,80 @@ library(caret)
 library(ISLR)
 library(tidyverse)
 library(caret)
+library(beepr)
 
 #load data
 ds <- read.csv("data/full_data_w_dummies_interaction.csv")
-ds <- subset(ds, select = -c(X, X.1) )
+ds$wrkstat <- as.factor(ds$wrkstat)
+ds$educcat <- as.factor(ds$educcat)
+ds$gender <- as.factor(ds$gender)
+ds$occrecode <- as.factor(ds$occrecode)
+ds$maritalcat <- as.factor(ds$maritalcat)
 
-#split in train, validation and test data
+# split in train, validation and test data
 set.seed(123)
 
-trainindex <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
-train <- ds[trainindex, ]
-remaining <- ds[-trainindex, ]
+trainindex_filter <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
+train <- ds[trainindex_filter, ]
+validation<- ds[-trainindex_filter, ]
 
-valid_test_index <- createDataPartition(y = remaining$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
-validation <- remaining[valid_test_index, ]
-test <- remaining[-valid_test_index, ]
+# valid_test_index_filter <- createDataPartition(y = remaining_filter$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
+# validation_filter <- remaining_filter[valid_test_index_filter, ]
+# test_filter <- remaining_filter[-valid_test_index_filter, ]
+
+cat("training dim: ", nrow(train), ", test dim: ", nrow(remaining), "\n")
+
+full.col.names <- colnames(ds)
+occrecode.labels <- full.col.names[match("armed_forces", full.col.names):match("transportation", full.col.names)]
+occrecode.labels <- occrecode.labels[occrecode.labels != "armed_forces"]
+educcat.labels <- full.col.names[match("bachelor", full.col.names):match("lessthan_high_school", full.col.names)]
+educcat.labels <- educcat.labels[educcat.labels != "junior_college"]
+prest.qt.labels <- c("prestg16until30", "prestg31until50", "prestggreater51")
[…]

### 2023-08-11 code/4_LassoRidgeElasticNet.R

@@ -10,36 +10,82 @@ library(ISLR)
 library(tidyverse)
 library(caret)
 library(glmnet)
+library(beepr)
 
 
 
 #load data 
 ds <- read.csv("data/full_data_w_dummies_interaction.csv")
-ds <- subset(ds, select = -c(X, X.1) )
+ds$wrkstat <- as.factor(ds$wrkstat)
+ds$educcat <- as.factor(ds$educcat)
+ds$gender <- as.factor(ds$gender)
+ds$occrecode <- as.factor(ds$occrecode)
+ds$maritalcat <- as.factor(ds$maritalcat)
 
-#split in train, validation and test data
+# split in train, validation and test data
 set.seed(123)
 
-trainindex <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
-train <- ds[trainindex, ]
-remaining <- ds[-trainindex, ]
+trainindex_filter <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
+train <- ds[trainindex_filter, ]
+validation<- ds[-trainindex_filter, ]
 
-valid_test_index <- createDataPartition(y = remaining$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
-validation <- remaining[valid_test_index, ]
-test <- remaining[-valid_test_index, ]
+# valid_test_index_filter <- createDataPartition(y = remaining_filter$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
+# validation_filter <- remaining_filter[valid_test_index_filter, ]
+# test_filter <- remaining_filter[-valid_test_index_filter, ]
+
+cat("training dim: ", nrow(train), ", test dim: ", nrow(validation), "\n")
+
+full.col.names <- colnames(ds)
+occrecode.labels <- full.col.names[match("armed_forces", full.col.names):match("transportation", full.col.names)]
+occrecode.labels <- occrecode.labels[occrecode.labels != "armed_forces"]
+educcat.labels <- full.col.names[match("bachelor", full.col.names):match("lessthan_high_school", full.col.names)]
+educcat.labels <- educcat.labels[educcat.labels != "junior_college"]
+prest.qt.labels <- c("prestg16until30", "prestg31until50", "prestggreater51")
[…]

### 2023-08-11 Paper/paper.Rmd
Updated code and paper

@@
 
-In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%), a validation (15%) and a test data set (15%).
+In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) dataset.
 We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RSME) using the validation data.
@@
 We remove all variables that have a higher than 90% correlation with each other.
-Our final dataset then contains 33,233 of 347 variables.
+Our final dataset then contains 33,244 observations of 361 variables.
+
+[Something about how we define the x variables for the models]
 
@@
 
-Our baseline linear model uses an OLS regression including all interaction variables. This results in an RMSE of 20,075.61.
+Our baseline linear model uses an OLS regression including all interaction variables. This results in an RMSE of 25,506.05.
 
-``` R
-mod_full <- train(realrinc ~  . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
+``` r
+formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+mod_full <- train(formula,
                   data = train, 
@@
 Given some extreme outliers in the income distribution of our data (which is typical for wage data) we follow the economic literature ([https://onlinelibrary.wiley.com/toc/14680084/2008/70/s1](#0)) [correct citation missing] and fit a logistic model to our data, using the log of `realrinc` as our dependent variable.
-This results in a higher RMSE than the linear model (35,844.61).
+This results in a higher RMSE than the linear model (25,909.82).
 
-``` R
-mod_loginc <- train(log_realrinc ~ . - occrecode -wrkstat - gender -educcat -maritalcat - age_sqr,
+``` r
+formula2 <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+mod_loginc <- train(formula2,
                     data = train, 
@@
 
-In both models the coefficient of female remains statistically insignificant at conventional levels. This suggests that gender is not a driving factor for wages within our data.
+In both models the coefficient of female is statistically significant with p \< 0.001.
+It suggests that being female decreases the real income by approximately 10,192\$ in the linear model or by 40% in the logistic model.
 
-## Support Vector Machines
+## Subset Selection
+
+As the dataset contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection.
+We use lasso, ridge and elastic net regressions.
+These methods allow us to fit a model that contains all variables.
+Then the model regularizes the coefficient estimates and shrinks them towards zero.
+In effect, we reduce the estimates' variance which improves the model fit.
+
+We start with a ridge regression.
+
+$$
+RSS + \lambda \sum^p _{j=1} \beta^2_j 
+$$
+
+Hence, the coefficient estimates in a ridge regression minimize the sum of the residual sum of squares (RSS) and the sum of squared coefficients multiplied with a tuning parameter $\lambda$.
+We estimate $\lambda$ using cross-validation.
+
+``` R
+ridge <- train(formula, data = train,
+  method = "glmnet", trControl = trainControl(method = "cv"),
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) 
+```
+
+Perform a ridge regression to predict real incomes yields an RMSE of 25,524.85.
+
+In contrast to other subset selection methods ridge regression does not exclude any variables from the regression model.
+Hence, the model is still fitted on 53 covariates.
+To overcome this issue we also use a lasso regression.
+
+$$
+RSS + \lambda \sum^p_{j=1} | \beta_j |
+$$
+
+In contrast to the ridge regression the lasso regression forces some of the coefficient estimates to be equal to zero if $\lambda$ is sufficiently large.
+Hence, the regression is not performed on all covariates.
+
+``` R
+lasso <- train(formula, data = train, 
+       method = "glmnet", trControl = trainControl(method = "cv"), 
+       tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
+```
+
+To perform the lasso regression in `R` we set $\alpha$ equal to one.
+The lasso regression yields a lower RMSE than the ridge regression (25,504.58 vs.
+25,524.85) and even a lower RMSE than the linear regression.
+
+In a final step we use elastic net regularization which is a linear combination of ridge and lasso regression.
+The elastic net regularization allows $\alpha$ to vary between zero and one.
+This results in an RMSE of 25,241.7.
+
+``` R
+elasticnet <- train(formula, data = train,
+       method = "glmnet", trControl = trainControl(method = "cv"),
+       tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
+       lambda = seq(from=0, to=0.15, by = 0.001)))
+```

### 2023-08-11 code/4_LassoRidgeElasticNet.R

@@ -76,24 +76,24 @@ x_cols_dummys <- x_cols_dummys[!x_cols_dummys %in% c("armed_forces", "other_wrks
 
 
 #Lasso
-lasso <- train(realrinc ~ female + . - log_realrinc, data = train, 
+formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+lasso <- train(formula, data = train, 
                 method = "glmnet", trControl = trainControl(method = "cv"), 
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
-caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25287.22
+caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.58
 
 
 #Ridge
-ridge <- train(realrinc ~ female + . - log_realrinc, data = train,
+ridge <- train(formula, data = train,
   method = "glmnet", trControl = trainControl(method = "cv"),
-  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #25320.77
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #25524.85
 caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
 
 #Elastic Net
-elasticnet <- train(realrinc ~ female + . - log_realrinc, data = train,
+elasticnet <- train(formula, data = train,
                     method = "glmnet", trControl = trainControl(method = "cv"),
                     tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
                     lambda = seq(from=0, to=0.15, by = 0.001)))
 caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #25241.7
 
 
-write.csv(ds, file = "data/final_data.csv")

### 2023-08-18 Paper/paper.Rmd
Update paper with references and literature review

@@
 In this paper we predict individuals' wages using data from the General Social Survey (GSS, 1974-2018).
-The variable of interest is `realrinc` which varies between 227\$ and 480,145\$.
+The variable of interest is `realrinc` which varies between 227\$ and 480,145\$\footnote{In constant 1986 USD.}.
 Hence, we can use regression techniques to predict individuals' wages.
@@
 
-To predict `realrinc` we use several regression techniques:
+The prediction of individuals' wages has been an issue in previous literature.
+As wages are relevant to companies they want to know how predict wages [@Chakraborti].
+Some papers make use of classification techniques to predict wages [@Chakraborti].
+As we have numeric values of real wages and not a categeorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests.
+We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler].
+Wages were already predicted using random forests by [@Eichinger].
+They find that usnig random forests results in better predictions than using linear models.
+Additionally, they show that gender barely effects the wage predictions.
+
+Based on the previous literature, we use severall regression techniques to predict `realrinc`:
 
@@
 
--   Support Vector Machines
+-   Lasso, Ridge and Elastic Net Regression
 
@@
 In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) dataset.
-We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RSME) using the validation data.
-
-[Write something about why to use RSME]
+We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RMSE) using the validation data.
+We choose the RMSE as it is more sensitive to large errors [@Arour].
 
@@
 [Something about ANOVA test for interaction variables].
+
+[Should we consider outlier handling? cf. Eichinger 2022]
 
@@
 
-Given some extreme outliers in the income distribution of our data (which is typical for wage data) we follow the economic literature ([https://onlinelibrary.wiley.com/toc/14680084/2008/70/s1](#0)) [correct citation missing] and fit a logistic model to our data, using the log of `realrinc` as our dependent variable.
+Given some extreme outliers in the income distribution of our data (which is typical for wage data) we follow the economic literature [@Ermini] and fit a logistic model to our data, using the log of `realrinc` as our dependent variable.
 This results in a higher RMSE than the linear model (25,909.82).
@@
 
-``` R
+``` r
 ridge <- train(formula, data = train,
@@
 
-``` R
+``` r
 lasso <- train(formula, data = train, 
@@
 
-``` R
+``` r
 elasticnet <- train(formula, data = train,

### 2023-08-21 code/5_models_GB_RF.R
add GB and RF script

@@ -1,0 +1,584 @@
+# set wd
+setwd("/home/user/Downloads/info_tud/statistical_learning_UDE/gender_pay_gap_project")
+# import functions file
+source('code/utils_functions.R')
+
+# packages
+library(caret)
+library(dplyr)
+library(lightgbm)
+library(Matrix)
+library(parallel)
+library(sgd)
+library(mlr3)
+library(mlr3learners)
+library(mlr3tuning)
+library(data.table)
+# remotes::install_github("mlr-org/mlr3extralearners@*release")
+library(mlr3extralearners)
+
+# load data
+ds <- fread("data/full_data_w_dummies_interaction.csv")
+ds$wrkstat <- as.factor(ds$wrkstat)
+ds$educcat <- as.factor(ds$educcat)
+ds$gender <- as.factor(ds$gender)
+ds$occrecode <- as.factor(ds$occrecode)
+ds$maritalcat <- as.factor(ds$maritalcat)
+
+# split in train, validation and test data
+set.seed(123)
+
+trainindex_filter <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
+train_filter <- ds[trainindex_filter, ]
+remaining_filter <- ds[-trainindex_filter, ]
+
+# write.csv(rbind(train_filter %>% 
+#   select(V1) %>% 
+#   mutate(label_ds = "TRAIN"),
+# remaining_filter %>% 
+#   select(V1) %>% 
+#   mutate(label_ds = "TEST")) %>% 
+#   arrange(V1,desc(TRUE)), "data/label_dataset.csv", row.names = FALSE)
+
+
+# valid_test_index_filter <- createDataPartition(y = remaining_filter$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
+# validation_filter <- remaining_filter[valid_test_index_filter, ]
+# test_filter <- remaining_filter[-valid_test_index_filter, ]
+
+cat("training dim: ", nrow(train_filter), ", test dim: ", nrow(remaining_filter), "\n")
+
+full.col.names <- colnames(ds)
+occrecode.labels <- full.col.names[match("armed_forces", full.col.names):match("transportation", full.col.names)]
+occrecode.labels <- occrecode.labels[occrecode.labels != "armed_forces"]
[…]

### 2023-08-22 code/utils_functions.R
data preparation explanation

@@ -13,3 +13,20 @@ estimate.vector <- function(i, ls_ds_interaction, all_dummys){
   result <- (all_dummys[,ls_ds_interaction[[i]][1]]*all_dummys[,ls_ds_interaction[[i]][2]])
   return(result)
 }
+
+simple_corr_matrix_plot <- function(df ,label_size, size_coef, title_graph, colnames_df, method ) {
+  df <- df %>% 
+    select_if(is.numeric)
+  #change colnames
+  colnames(df)<- colnames_df
+  #correlation matrix
+  corr_matrix <- cor(df, method = method)
+  corrplot(corr_matrix , type="lower",
+           tl.cex= label_size,
+           tl.offset = 0.5, tl.col = "black", method = "color",
+           number.cex= 1, tl.srt = 45,
+           addCoef.col = "black",
+           diag = FALSE,
+           title = title_graph, outline=FALSE, cl.pos="b",
+           number.digits=2, mar=c(0,0,0,0))
+}

### 2023-08-22 Paper/paper.Rmd

@@
 
-# Data Preparation
+# Data Preparation & Feature Engineering
 
-Before conducting the regression analysis we prepared the data.
-The provided dataset has 54,842 observations and for 38.55% of the observations the outcome variable `realrinc` is missing.
-As we cannot predict the wage for those observations we removed them.
-This left us with a dataset containing 33,702 observations.
-Additionally, we do not know the occupation for some observations (`occ10 = 9997` ) and removed those.
-This diminishes the dataset to 33,244 observations.
+The initial data set contains 11 variables and a brief description of each variable is presented here:
 
-In this dataset still there were missing some values for some observations, e.g., number of children or occupation.
-To impute these values we used the `mice` function.
-We imputed age, occupational prestige and number of children with predictive mean matching.
-For the factor variables education and marital status we used polytomous logisitic regression.
-We receive six imputation datasets and use mean estimation for age, occupational prestige and number of children.
+- **year**: survey year
+- **realrinc**: respondent's base income (in constant 1986 USD)
+- **age**: respondent's age in years
+- **occ10**: respondent's occupation code
+- **occrecode**: recode of the occupation code into one of 11 main categories
+- **prestg10**: respondent's occupational prestige score
+- **childs**: number of children
+- **wrkstat**: work status of the respondent (full-time, part-time, temporarily not working, unemployed (laid off), etc)
+- **gender**: respondent's gender
+- **educcat**: respondent's degree level (Less Than High School, High School, etc)
+- **maritalcat**: respondent's marital status (Married, Widowed, Divorced, etc)
 
-With this dataset we create dummy variables for the factor variables. Additionally, we create interaction variables.
+## Data Preparation
 
-[Something about ANOVA test for interaction variables].
+A data preparation process was carried before the modeling.
+The initial dataset had 54,842 observations and the outcome variable `realrinc` had 38.55% of missing values.
+The imputation of the target variable was discarded to avoid generating a bias in the modeling and therefore it was decided to eliminate these records from the original data set.
+After this, the dataset contained 33,702 observations.
+Furthermore, it was impossible to deduce the occupation for some observations (`occ10 = 9997`) and these were removed from the dataset.
+At the end, the final dataset had 33,244 observations, and it was used for the modeling process.
 
-[Should we consider outlier handling? cf. Eichinger 2022]
+This dataset still had some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others.
+The imputation process was carried using five variables via `mice` package.
+The features *age, occupational prestige and number of children* were imputed using the pmm (`predictive mean matching`) method, and whereas the polytomous logistic regression method was used for the factor variables (*education and marital status*).
+The imputation process used a number of multiple imputations of six, which means that the process had 6 different datasets, and then the estimation of the mean and the majority vote was applied to aggregate the datasets into a single dataset.
+The mean estimation was applied for the numeric variables and majority vote for the categorical variables..
 
-In a final step we calculate the correlation matrix using Spearman's rank correlation coefficient.
-We remove all variables that have a higher than 90% correlation with each other.
-Our final dataset then contains 33,244 observations of 361 variables.
+## Feature Engineering
 
-[Something about how we define the x variables for the models]
+The interaction variables were created after the imputation process.
+First, some interaction features were estimated using only the numerical variables, and then the numerical variables were transformed into categorical variables by dividing them into different ranges (e.g.
+age between 18 and 30, age between 31 and 50, etc).
+These new categorical variables and the initial categorical variables (*marital status, education, etc*) were transformed into dummy variables.
+And finally, the interaction variables were estimated with the group of dummy variables mentioned above.
+The inclusion of the interaction of the numeric variables were validate using a linear model, and the results showed that the p- values of the parameters for the interactions were less than 0,05.
+
+![table](includes/reg_table.png)
+
+Furthermore, a correlation matrix using Spearman's rank correlation coefficient was estimated with some numeric variables in order to support the previous results.
+The correlation results are shown in the following figure.
+
+![matrix_corr](includes/corr_matrix.png)
+
+After the data preparation and feature engineering processes, a final dataset was constructed with 33,244 observations and 421 features.
+Due to the large number of interaction variables, it was decided to reduce them and use only the interaction between a group of variables (*occrecode, educcat and prestg10*).
+After reduction, the modeling process was developed with 100 variables (numerical and interaction variables).
+The following figure shows the overview of the group of variables that were estimated and used in the models.
+
+![feature_process](includes/feature_process.png)
 

### 2023-08-22 Paper/paper.tex

\documentclass[11pt,a4paper]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{???}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm]{geometry}

% Figure Placement:
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

%% citation setup
\usepackage{csquotes}

\usepackage[backend=biber, maxbibnames = 99, style = apa]{biblatex}
\setlength\bibitemsep{1.5\itemsep}
\addbibresource{R_packages.bib}
\bibliography{references.bib}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true, linktocpage = TRUE]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Alex Amaguaya, Lea Bergmann},
            pdftitle={The Gender Pay Gap in the General Social Survey},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{The Gender Pay Gap in the General Social Survey}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Statistical Learning}
  \author{Alex Amaguaya, Lea Bergmann}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{today}

%% linespread settings

\usepackage{setspace}

\onehalfspacing

% Language Setup

\usepackage{ifthen}
\usepackage{iflang}
\usepackage[super]{nth}
\usepackage[ngerman, english]{babel}

%Acronyms
\usepackage[printonlyused, withpage, nohyperlinks]{acronym}
\usepackage{changepage}

% Multicols for the Title page
\usepackage{multicol}

\usepackage{longtable}

\begin{document}

\selectlanguage{english}

%\maketitle

\begin{titlepage}
  \noindent\begin{minipage}{0.6\textwidth}
	  \IfLanguageName{english}{University of Duisburg-Essen}{Universit\"at Duisburg-Essen}\\
	  \IfLanguageName{english}{Faculty of Business Administration and Economics}{Fakult\"at f\"ur Wirtschaftswissensschaften}\\
	  \IfLanguageName{english}{Chair of Econometrics}{Lehrstuhl f\"ur \"Okonometrie}\\
  \end{minipage}
	\begin{minipage}{0.4\textwidth}
	  \begin{flushright}
  	  \vspace{-0.5cm}
      \IfLanguageName{english}{\includegraphics*[width=5cm]{Includes/duelogo_en.png}}{\includegraphics*[width=5cm]{Includes/duelogo_de.png}}
	  \end{flushright}
	\end{minipage}
  \\
  \vspace{1.5cm}
  \begin{center}
  \huge{The Gender Pay Gap in the General Social Survey}\\
  \vspace{.25cm}
  \Large{Statistical Learning}\\
  \vspace{0.5cm}
  \large{Term Paper}\\
  \vspace{1cm}
  \large{
  \IfLanguageName{english}{Submitted to the Faculty of \\ Business Administration and Economics \\at the \\University of Duisburg-Essen}{Vorgelegt der \\Fakult\"at f\"ur Wirtschaftswissenschaften der \\ Universit\"at Duisburg-Essen}\\}
  \vspace{0.75cm}
  \large{\IfLanguageName{english}{from:}{von:}}\\
  \vspace{0.5cm}
  Alex Amaguaya, Lea Bergmann\\
  \end{center}
  %\vspace{2cm}
  \vfill
  \hrulefill

  \noindent\begin{minipage}[t]{0.3\textwidth}
  \IfLanguageName{english}{Reviewer:}{Erstgutachter:}
  \end{minipage}
  \begin{minipage}[t]{0.7\textwidth}
  \hspace{1cm}
  \end{minipage}

  \noindent\begin{minipage}[t]{0.3\textwidth}
  \IfLanguageName{english}{Deadline:}{Abgabefrist:}
  \end{minipage}
  \begin{minipage}[t]{0.7\textwidth}
  \hspace{1cm}tomorrow
  \end{minipage}

  \hrulefill

  \begin{multicols}{3}

  Name:

  Matriculation No.:

  E-Mail:

  Study Path:

  Semester:

  Graduation (est.):

  \columnbreak

  Ape Monkey

  123456

  Ape@Monkey.biz

  M.Sc. Economics

  \nth{5}

  Summer Term 2020

  \columnbreak

  John Doe

  234567

  john.doe@web.de

  M.Sc. Economics

  \nth{4}

  Summer Term 2020

  \columnbreak

  Darth Vader

  543556

  vader@emperialenterprises.com

  M.Sc. Economics

  \nth{3}

  Summer Term 2020

  \end{multicols}

\end{titlepage}

\newgeometry{top=2cm, left = 5cm, right = 2.5cm, bottom = 2.5cm}

\pagenumbering{Roman}
{
\hypersetup{linkcolor=black}

\setcounter{tocdepth}{3}
\tableofcontents
}

\newpage
\listoffigures
\addcontentsline{toc}{section}{List of Figures}

%\newpage
\listoftables
\addcontentsline{toc}{section}{List of Tables}

\section*{List of Abbreviations}
\addcontentsline{toc}{section}{List of Abbreviations}

\begin{adjustwidth}{1.5em}{0pt}

\begin{acronym}[dummyyyy]
 \acro{ECTSCP}{European Credit Transfer System Credit Point}
 \acro{lasso}{Least Absolute Shrinkage and Selection Operator}
 \acro{pcr}{Principal Components Regression}
 \acro{pls}{Partial Least Squares}
 \acro{RMSE}{Root Mean Squared Error}
 \acroplural{LRG}[LRG]{laengefristige Refinanzierungsgeschaefte}

%Falls eine Abkuerzung in der Mehrzahl nicht einfach auf "s" endet muss das speziell eingestellt werden.
% \acro{slmtA}{super lange mega tolle Abkuerzung} %Einzahl
 %\acroplural{slmtA}[slmtAs]{super lange mega tolle Abkuerzungen} %Mehrzahl
 \acro{dummyyyy}{dummyyy}
\end{acronym}

\end{adjustwidth}

\restoregeometry

\newpage
\pagenumbering{arabic}
\hypertarget{rmarkdown-template}{%
\section{Rmarkdown Template}\label{rmarkdown-template}}

Currently there is one thing you need to customise manually in
\emph{template.tex} which will be used by the \LaTeX processor for
generating the PDF: the entries in the columns on the title page
containing student info. Here you have to replace the dummy data:

\begin{verbatim}
\begin{multicols}{$cols_authors$}

Name:

Matriculation No.:

E-Mail:

Study Path:

Semester:

Graduation (est.):
\end{verbatim}

If you are just two people, working alone or want to remove the colum
design you may delete (parts of) this section.

\hypertarget{r-markdown}{%
\subsection{R Markdown}\label{r-markdown}}

This is an R Markdown document. Markdown is a simple formatting syntax
for authoring HTML, PDF, and MS Word documents. For more details see
\url{http://rmarkdown.rstudio.com}.

You may use \LaTeX~to write formulas, e.g., \(X^2 = \sqrt{X^4}\) and
\[X^2 = \sqrt{X^4}.\]

After clicking the \emph{Knit} button in \emph{RStudio}, a (PDF)
document that includes both content and the output of any embedded R
code chunks will be generated.

You can embed an R code chunk like this:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{summary}\NormalTok{(cars)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      speed           dist       
##  Min.   : 4.0   Min.   :  2.00  
##  1st Qu.:12.0   1st Qu.: 26.00  
##  Median :15.0   Median : 36.00  
##  Mean   :15.4   Mean   : 42.98  
##  3rd Qu.:19.0   3rd Qu.: 56.00  
##  Max.   :25.0   Max.   :120.00
\end{verbatim}

\hypertarget{including-plots}{%
\subsection{Including Plots}\label{including-plots}}

You can also embed plots, for example:

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/pressure-1.pdf}
\caption{\label{fig:pressure} Pressure}
\end{figure}

Note that the \texttt{echo\ =\ FALSE} parameter was added to the code
chunk to prevent printing of the R code that generated the plot. You can
label the plot above by including the label
\texttt{\textbackslash{}\textbackslash{}label\{fig:pressure\}} in the
chunk argument \texttt{fig.cap}. A reference to the plot is then made as
follows:

Looking at Figure \ref{fig:pressure} makes me happy.

\hypertarget{the-yaml-header}{%
\subsection{The YAML Header}\label{the-yaml-header}}

The YAML header is at the very top of this document. It is enclosed by 3
dashes. Although some useful options are specified already (most of
which are hopefully self-explaining) you may want additional
customisation. Below we discuss useful options that we have implemented
for you.

\emph{Note that \textbf{indentation and line breaks matter} in the YAML
header.} More about that
\href{https://learn-the-web.algonquindesign.ca/topics/markdown-yaml-cheat-sheet/\#yaml}{here}.

\hypertarget{language}{%
\subsubsection{\texorpdfstring{\texttt{language}}{language}}\label{language}}

This variable affects mainly the headings of your output file and
automatic hyphenation.

You can set this to \texttt{english} or \texttt{german} like this:

\texttt{language:\ english}

\texttt{language:\ german}

If the language is not specified in the YAML header it will be set to
\texttt{english}.

\hypertarget{linespread}{%
\subsubsection{\texorpdfstring{\texttt{linespread}}{linespread}}\label{linespread}}

The default value for the linespread is 1.5. Usually this is fine and
sometimes it's required. If you nevertheless want to change it you can
do so by specifying the linespread variable, e.g.

\texttt{linespread:\ 1}

\hypertarget{rmarkdown-makes-your-life-easy}{%
\section{Rmarkdown makes your life
easy}\label{rmarkdown-makes-your-life-easy}}

\hypertarget{the-kable-function}{%
\subsection{The kable function}\label{the-kable-function}}

In empirical work it's crucial not only to present your results but also
to explain your research strategy. This often involves tables presenting
data and results. Generating tables by hand using \LaTeX is an option
but may be time consuming. However, there is a variety of R packages
that automate this. One of them is the \texttt{kable} package. It can
generate \LaTeX tables from a variety of R Objects.

Example: you are working on an analysis of black cherry trees and want
to present \(n\) observations to the reader You can do that using
\texttt{knitr::kable()}.

\begin{longtable}[]{@{}rrr@{}}
\caption{6 Observations from the trees Dataset}\tabularnewline
\toprule\noalign{}
Diameter & Height & Volume \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Diameter & Height & Volume \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
17.9 & 80 & 58.3 \\
8.3 & 70 & 10.3 \\
10.8 & 83 & 19.7 \\
10.7 & 81 & 18.8 \\
18.0 & 80 & 51.0 \\
8.6 & 65 & 10.3 \\
\end{longtable}

Now that we have presented our data it's analysis time! Lets start with
a quick call to summary().

\hypertarget{the-stargazer-package}{%
\subsection{The Stargazer package}\label{the-stargazer-package}}

Calling summary() in a code chunk will work but this will give you quite
an ugly result (just try it for yourself!). When it comes to presenting
more structured objects like summaries, model results or for example
correlation matrices the stargazer package is well suited.

\begin{table}[!htbp] \centering 
  \caption{Summary} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Girth & 31 & 13.248 & 3.138 & 8.300 & 20.600 \\ 
Height & 31 & 76.000 & 6.372 & 63 & 87 \\ 
Volume & 31 & 30.171 & 16.438 & 10.200 & 77.000 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

Assume we want to evaluate how the height and the volume of a typical
cherry tree are related. We are estimating this using OLS to estimate a
simple linear model.

\begin{table}[!htbp] \centering 
  \caption{Regression results} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Volume \\ 
\hline \\[-1.8ex] 
 Height & 1.543$^{***}$ \\ 
  & (0.384) \\ 
  & \\ 
 Constant & $-$87.124$^{***}$ \\ 
  & (29.273) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 31 \\ 
R$^{2}$ & 0.358 \\ 
Adjusted R$^{2}$ & 0.336 \\ 
Residual Std. Error & 13.397 (df = 29) \\ 
F Statistic & 16.164$^{***}$ (df = 1; 29) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

Now that we have our model we can visualize it.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Dataset and regression}
\end{figure}

By the way: the (\LaTeX) command \texttt{\textbackslash{}pagebreak} can
be used to force a page break.

\pagebreak

\hypertarget{citations}{%
\subsection{Citations}\label{citations}}

A \emph{bibtex} bibliography can be used for citations. The bibliography
file used in this template is \texttt{references.bib} and you find it in
the project directory. It is easily edited using \emph{RStudio}

A simple \emph{bibtex} entry looks like this:

\begin{verbatim}
@book{
  Hastie2013,
  publisher = {Springer},
  year = {2013},
  title = {The elements of statistical learning},
  author = {Hastie, Trevor and Tibshirani, Robert and Friedman, Jerome H.},
}
\end{verbatim}

The first field (\texttt{Hastie2013}) is the identifier which allows you
to cite a reference.

You can cite a source in Harvard style like this: \autocite{Hastie2013}
or \textcite{Hastie2013}.

A cited source will be automatically added to the reference section at
the end of the document.

\pagebreak

\hypertarget{conclusion}{%
\section{\texorpdfstring{Conclusion
\label{chap:conc}}{Conclusion }}\label{conclusion}}

Your conclusion. Note that you may also refer to a specific chapter or
section in the document, provided there is a label. We have anchored a
label to the section header of the Conclusion. You may reference it as
follow: See Chapter \ref{chap:conc}.

This section allows you to reference R packages used in the analysis.
Simply include them in \texttt{R\_packages.bib} as \emph{bibtex} entries
and include the identifiers using
\texttt{\textbackslash{}notecite\{...\}} as shown below.

\pagebreak

\addcontentsline{toc}{section}{References}
\printbibliography[title = References]
\cleardoublepage

\begin{refsection}
\nocite{R-base}
\nocite{R-Studio}

\printbibliography[title = Software-References]
\addcontentsline{toc}{section}{Software-References}
\end{refsection}

\cleardoublepage
\appendix
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}

\newgeometry{top=2.5cm, left = 2cm, right = 2cm, bottom = 2.5cm}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{description-of-relevant-variables}{%
\subsection{\texorpdfstring{Description of relevant Variables
\label{app:A}}{Description of relevant Variables }}\label{description-of-relevant-variables}}

An example using the LaTeX environment \emph{longtable}:

\begin{longtable}{l|l|p{6cm}}
\caption{Description of relevant variables} \\
\hline
\textbf{Variable} & \textbf{Dataset} & \textbf{Description}    \\
\hline
\endfirsthead % Erster Kopf zu Ende
% Definition des Tabellenkopfes auf den folgenden Seiten
\caption{Description of relevant variables (continued)}\\
\hline
\textbf{Variable} & \textbf{Dataset} & \textbf{Description}    \\
\hline
\endhead % Zweiter Kopf ist zu Ende
\multicolumn{3}{r}{Continued on the next page.}\\
\endfoot
\hline
\multicolumn{3}{r}{End of table.} \\
\endlastfoot
Student.Pseudonym & all & Student’s immatriculation number \\
\hline
Semester & FS Data & Academic semester. The first four digits stand for the year, the fifth digit is either a 1 (summer semester) or a 2 (winter semester). Example: "20062" describes the winter semester 2006/2007, "20071" the summer semester 2007 \\
\hline
FS & FS, Pruefung Data & Study semester \\
\hline
Status & FS Data & Status as a student. "R" stands for re-enrollment, "N" for new enrollment, "E" for initial enrollment and "B" for leave of absence \\
\hline
Status.dpp & Pruefung Data & See variable Status \\
\hline
Austrittsgrund & FS, Studium Data & Reason for dropping out \\
\hline
SGCode & all & Code for study program. The last digit represents the examination regulations (PO). A change to the PO is thus considered as a new program in the system \\
\hline
Abschluss\_Bezeichnung & FS, Pruefung Data & Name of the degree that students receive upon successful completion of their studies \\
\hline
Fach\_Bezeichnung & FS, Pruefung Data & Name of the course of studies \\
\hline
PO & all & Examination regulations \\
\hline
PolyvalentePruefungsnummer & Pruefung Data & Polyvalent examination number \\
\hline
Bezeichnung & Pruefung Data & Name of the examination \\
\hline
Pruefungssemester & Pruefung Data & Semester in which the student was registered for the exam. Structure analogous to the variable ‘Semester’ in FS Data \\
\hline
Status.Pruefung & Pruefung Data & Status of the exam. "BE" means Passed, "ZU" means Withdrawn, "NB" means Not Passed and "PV" means Examination Existing (applications or performances are available, but the whole module is not yet completed) \\
\hline
Versuchszahl & Pruefung Data & Trial number \\
\hline
Note & Pruefung Data & Grade \\
\hline
VerbuchteECTSCP & Pruefung Data & Received ECTSCP \\
\hline
Start\_Semester & Studium Data & Semester in which the studies were started. Structure analogous to the variable ‘Semester’ in FS Data \\
\hline
Ende\_Semester & Studium Data & Semester in which the studies were finshed. Structure analogous to the variable ‘Semester’ in FS Data
\label{tab_var}
\end{longtable}

\end{document}

### 2023-08-23 Paper/paper.tex
Minor changes in the paper

@@
 \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
 \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
@@
 \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
 \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
@@
 \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
 \newcommand{\ImportTok}[1]{#1}
@@
 \newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
 \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{longtable,booktabs}
 \usepackage{graphicx}
@@
 \pagenumbering{arabic}
-\hypertarget{rmarkdown-template}{%
-\section{Rmarkdown Template}\label{rmarkdown-template}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Currently there is one thing you need to customise manually in
-\emph{template.tex} which will be used by the \LaTeX processor for
-generating the PDF: the entries in the columns on the title page
-containing student info. Here you have to replace the dummy data:
+In this paper we predict individuals' wages using data from the General
+Social Survey (GSS, 1974-2018). The variable of interest is
+\texttt{realrinc} which varies between 227\$ and
+480,145\$\footnote{In constant 1986 USD.}. Hence, we can use regression
+techniques to predict individuals' wages. Additionally, we can consider
+whether \texttt{realrinc} varies by gender and if so, how much.
 
-\begin{verbatim}
-\begin{multicols}{$cols_authors$}
+The data from the GSS allow us to consider various control variables
+such as gender, age, education, occupation, number of children and
+marital status.
 
-Name:
+The prediction of individuals' wages has been an issue in previous
+literature. As wages are relevant to companies they want to know how
+predict wages \autocite{Chakraborti}. Some papers make use of
+classification techniques to predict wages \autocite{Chakraborti}. As we
+have numeric values of real wages and not a categorical variable we do
+not make use of classification techniques but prefer linear regression
+techniques such as OLS and Lasso regressions as well as random forests.
+We prefer these techniques as they handle categorical variables and
+outliers well, which are features of our data \autocite{Cutler}. Wages
+were already predicted using random forests by \autocite{Eichinger}.
+They find that using random forests results in better predictions than
+using linear models. Additionally, they show that gender barely effects
+the wage predictions.
 
-Matriculation No.:
+Based on the previous literature, we use several regression techniques
+to predict \texttt{realrinc}:
 
-E-Mail:
+\begin{itemize}
+\item
+  Linear Regression
+\item
+  Logistic Regression
+\item
+  Lasso, Ridge and Elastic Net Regression
+\item
+  Gradient Boosting
+\item
+  Random Forest
+\end{itemize}
 
-Study Path:
+In order to evaluate which of the regression techniques predicts
+\texttt{realrinc} the best we split the data set into a train (70\%) and
+a test (30\%) data set. We fit the aforementioned regression models on
+the train data and evaluate the root mean squared error (RMSE) using the
+validation data. We choose the RMSE as it is more sensitive to large
+errors \autocite{Arour}.
 
-Semester:
+{[}Write something about cross validation?{]}
 
-Graduation (est.):
-\end{verbatim}
+\hypertarget{data-preparation-feature-engineering}{%
+\section{Data Preparation \& Feature
+Engineering}\label{data-preparation-feature-engineering}}
 
-If you are just two people, working alone or want to remove the colum
-design you may delete (parts of) this section.
+The initial data set contains 11 variables, which are described in the
+following: - \textbf{year}: survey year - \textbf{realrinc}:
+respondent's base income (in constant 1986 USD) - \textbf{age}:
+respondent's age in years - \textbf{occ10}: respondent's occupation code
+- \textbf{occrecode}: recode of the occupation code into one of 11 main
+categories - \textbf{prestg10}: respondent's occupational prestige score
+- \textbf{childs}: number of children - \textbf{wrkstat}: work status of
+the respondent (full-time, part-time, temporarily not working,
+unemployed (laid off), etc.) - \textbf{gender}: respondent's gender -
+\textbf{educcat}: respondent's degree level (Less Than High School, High
+School, etc.) - \textbf{maritalcat}: respondent's marital status
+(Married, Widowed, Divorced, etc.)
 
-\hypertarget{r-markdown}{%
-\subsection{R Markdown}\label{r-markdown}}
+\hypertarget{data-preparation}{%
+\subsection{Data Preparation}\label{data-preparation}}
 
-This is an R Markdown document. Markdown is a simple formatting syntax
-for authoring HTML, PDF, and MS Word documents. For more details see
-\url{http://rmarkdown.rstudio.com}.
+Before modeling we prepared the data. The initial data set has 54,842
+observations and the outcome variable \texttt{realrinc} has 38.55\% of
+missing values. The imputation of the target variable was discarded to
+avoid generating a bias in the modeling and therefore we decided to
+eliminate these records from the original data set. After this, the data
+set contains 33,702 observations. Furthermore, it was impossible to
+deduce the occupation for some observations (\texttt{occ10\ =\ 9997})
+and these are removed from the data set. At the end, the final data set
+has 33,244 observations, and is used for the modeling process.
 
-You may use \LaTeX~to write formulas, e.g., \(X^2 = \sqrt{X^4}\) and
-\[X^2 = \sqrt{X^4}.\]
+This data set still has some missing values for the regressors, e.g.,
+number of children, age, marital status, degree level, and others. Using
+the \texttt{mice} package we conduct the imputation using five
+variables. We impute the features \emph{age, occupational prestige and
+number of children} using the pmm (\texttt{predictive\ mean\ matching})
+method. The polytomous logistic regression method is used for the factor
+variables (\emph{education and marital status}). The imputation process
+uses a number of multiple imputations of six. This means that the
+imputation process creates 6 different data sets, and then the
+estimation of the mean and the majority vote is applied to aggregate the
+data sets into a single data set. The mean estimation is applied for
+numeric variables and majority vote for categorical variables..
 
-After clicking the \emph{Knit} button in \emph{RStudio}, a (PDF)
-document that includes both content and the output of any embedded R
-code chunks will be generated.
+\hypertarget{feature-engineering}{%
+\subsection{Feature Engineering}\label{feature-engineering}}
 
-You can embed an R code chunk like this:
+After the imputation process we create interaction variables. First,
+some interaction features are estimated using only the numerical
+variables, and then we transform the numerical variables into
+categorical variables by dividing them into different ranges (e.g.~age
+between 18 and 30, age between 31 and 50, etc). These new categorical
+variables and the initial categorical variables (\emph{marital status,
+education, etc}) are transformed into dummy variables. And finally, we
+estimate the interaction variables with the group of dummy variables
+mentioned above. We validate the inclusion of the interaction of the
+numeric variables using a linear model. The results show that the
+p-values of the parameters for the interactions are less than 0.05.
+
+\begin{figure}
+\centering
+\includegraphics{includes/reg_table.png}
+\caption{table}
+\end{figure}
+
+Furthermore, we estimate a correlation matrix using Spearman's rank
+correlation coefficient with some numeric variables in order to support
+the previous results. The correlation results are shown in the following
+figure.
+
+\begin{figure}
+\centering
+\includegraphics{includes/corr_matrix.png}
+\caption{matrix\_corr}
+\end{figure}
+
+After the data preparation and feature engineering processes, we
+construct a final data set with 33,244 observations and 421 features.
+Due to the large number of interaction variables, we decided to reduce
+them and use only the interaction between a group of variables
+(\emph{occrecode, educcat and prestg10}). After reduction, the modeling
+process is developed with 100 variables (numerical and interaction
+variables). The following figure shows the overview of the group of
+variables that we estimated and used in the models.
+
+\begin{figure}
+\centering
+\includegraphics{includes/feature_process.png}
+\caption{feature\_process}
+\end{figure}
+
+\hypertarget{model-training-and-tuning}{%
+\section{Model Training and Tuning}\label{model-training-and-tuning}}
+
+We use five different regression techniques to predict the individuals'
+wages using the final dataset. We fit every model on our training data
+using the \texttt{train} function from the \texttt{caret} package. To
+choose the best parameter combination we use cross-validation within all
+train functions.
+
+\hypertarget{linear-and-logistic-regression}{%
+\subsection{Linear and Logistic
+Regression}\label{linear-and-logistic-regression}}
+
+Our baseline linear model uses an OLS regression including all
+interaction variables. This results in an RMSE of 25,506.05.
 
@@
 \begin{Highlighting}[]
-\FunctionTok{summary}\NormalTok{(cars)}
+\NormalTok{formula }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"realrinc \textasciitilde{}"}\NormalTok{, }\FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
+\NormalTok{mod\_full }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula,}
+                  \AttributeTok{data =}\NormalTok{ train, }
+                  \AttributeTok{method =} \StringTok{"lm"}\NormalTok{,  }
+                  \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{))}
 \end{Highlighting}
@@
 
-\begin{verbatim}
-##      speed           dist       
-##  Min.   : 4.0   Min.   :  2.00  
-##  1st Qu.:12.0   1st Qu.: 26.00  
-##  Median :15.0   Median : 36.00  
-##  Mean   :15.4   Mean   : 42.98  
-##  3rd Qu.:19.0   3rd Qu.: 56.00  
-##  Max.   :25.0   Max.   :120.00
-\end{verbatim}
+Given some extreme outliers in the income distribution of our data
+(which is typical for wage data) we follow the economic literature
+\autocite{Ermini} and fit a logistic model to our data, using the log of
+\texttt{realrinc} as our dependent variable. This results in a higher
+RMSE than the linear model (25,909.82).
 
-\hypertarget{including-plots}{%
-\subsection{Including Plots}\label{including-plots}}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{formula2 }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"log\_realrinc \textasciitilde{}"}\NormalTok{, }\FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
+\NormalTok{mod\_loginc }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula2,}
+                    \AttributeTok{data =}\NormalTok{ train, }
+                    \AttributeTok{method =} \StringTok{"lm"}\NormalTok{,  }
+                    \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
-You can also embed plots, for example:
+In both models the coefficient of female is statistically significant
+with p \textless{} 0.001. It suggests that being female decreases the
+real income by approximately 10,192\$ in the linear model or by 40\% in
+the logistic model.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/pressure-1.pdf}
-\caption{\label{fig:pressure} Pressure}
-\end{figure}
+\hypertarget{subset-selection}{%
+\subsection{Subset Selection}\label{subset-selection}}
 
-Note that the \texttt{echo\ =\ FALSE} parameter was added to the code
-chunk to prevent printing of the R code that generated the plot. You can
-label the plot above by including the label
-\texttt{\textbackslash{}\textbackslash{}label\{fig:pressure\}} in the
-chunk argument \texttt{fig.cap}. A reference to the plot is then made as
-follows:
+As the dataset contains many variables (361) and even our selection of
+dummy variables for the right-hand side of the regression leaves us with
+53 variables, we make use of regression techniques for subset selection.
+We use lasso, ridge and elastic net regressions. These methods allow us
+to fit a model that contains all variables. Then the model regularizes
+the coefficient estimates and shrinks them towards zero. In effect, we
+reduce the estimates' variance which improves the model fit.
 
-Looking at Figure \ref{fig:pressure} makes me happy.
+We start with a ridge regression.
 
-\hypertarget{the-yaml-header}{%
-\subsection{The YAML Header}\label{the-yaml-header}}
+\[
+RSS + \lambda \sum^p _{j=1} \beta^2_j 
+\]
 
-The YAML header is at the very top of this document. It is enclosed by 3
-dashes. Although some useful options are specified already (most of
-which are hopefully self-explaining) you may want additional
-customisation. Below we discuss useful options that we have implemented
-for you.
+Hence, the coefficient estimates in a ridge regression minimize the sum
+of the residual sum of squares (RSS) and the sum of squared coefficients
+multiplied with a tuning parameter \(\lambda\). We estimate \(\lambda\)
+using cross-validation.
 
-\emph{Note that \textbf{indentation and line breaks matter} in the YAML
-header.} More about that
-\href{https://learn-the-web.algonquindesign.ca/topics/markdown-yaml-cheat-sheet/\#yaml}{here}.
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ridge }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula, }\AttributeTok{data =}\NormalTok{ train,}
+  \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{),}
+  \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{0}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{500}\NormalTok{, }\DecValTok{1}\NormalTok{))) }
+\end{Highlighting}
+\end{Shaded}
 
-\hypertarget{language}{%
-\subsubsection{\texorpdfstring{\texttt{language}}{language}}\label{language}}
+Perform a ridge regression to predict real incomes yields an RMSE of
+25,524.85.
 
-This variable affects mainly the headings of your output file and
-automatic hyphenation.
+In contrast to other subset selection methods ridge regression does not
+exclude any variables from the regression model. Hence, the model is
+still fitted on 53 covariates. To overcome this issue we also use a
+lasso regression.
 
-You can set this to \texttt{english} or \texttt{german} like this:
+\[
+RSS + \lambda \sum^p_{j=1} | \beta_j |
+\]
 
-\texttt{language:\ english}
+In contrast to the ridge regression the lasso regression forces some of
+the coefficient estimates to be equal to zero if \(\lambda\) is
+sufficiently large. Hence, the regression is not performed on all
+covariates.
 
-\texttt{language:\ german}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{lasso }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula, }\AttributeTok{data =}\NormalTok{ train, }
+       \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{), }
+       \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{1}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{500}\NormalTok{,}\DecValTok{1}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
 
-If the language is not specified in the YAML header it will be set to
-\texttt{english}.
+To perform the lasso regression in \texttt{R} we set \(\alpha\) equal to
+one. The lasso regression yields a lower RMSE than the ridge regression
+(25,504.58 vs.~25,524.85) and even a lower RMSE than the linear
+regression.
 
-\hypertarget{linespread}{%
-\subsubsection{\texorpdfstring{\texttt{linespread}}{linespread}}\label{linespread}}
+In a final step we use elastic net regularization which is a linear
+combination of ridge and lasso regression. The elastic net
+regularization allows \(\alpha\) to vary between zero and one. This
+results in an RMSE of 25,241.7.
 
-The default value for the linespread is 1.5. Usually this is fine and
-sometimes it's required. If you nevertheless want to change it you can
-do so by specifying the linespread variable, e.g.
-
-\texttt{linespread:\ 1}
-
-\hypertarget{rmarkdown-makes-your-life-easy}{%
-\section{Rmarkdown makes your life
-easy}\label{rmarkdown-makes-your-life-easy}}
-
-\hypertarget{the-kable-function}{%
-\subsection{The kable function}\label{the-kable-function}}
-
-In empirical work it's crucial not only to present your results but also
-to explain your research strategy. This often involves tables presenting
-data and results. Generating tables by hand using \LaTeX is an option
-but may be time consuming. However, there is a variety of R packages
-that automate this. One of them is the \texttt{kable} package. It can
-generate \LaTeX tables from a variety of R Objects.
-
-Example: you are working on an analysis of black cherry trees and want
-to present \(n\) observations to the reader You can do that using
-\texttt{knitr::kable()}.
-
-\begin{longtable}[]{@{}rrr@{}}
-\caption{6 Observations from the trees Dataset}\tabularnewline
-\toprule\noalign{}
-Diameter & Height & Volume \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-Diameter & Height & Volume \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-17.9 & 80 & 58.3 \\
-8.3 & 70 & 10.3 \\
-10.8 & 83 & 19.7 \\
-10.7 & 81 & 18.8 \\
-18.0 & 80 & 51.0 \\
-8.6 & 65 & 10.3 \\
-\end{longtable}
-
-Now that we have presented our data it's analysis time! Lets start with
-a quick call to summary().
-
-\hypertarget{the-stargazer-package}{%
-\subsection{The Stargazer package}\label{the-stargazer-package}}
-
-Calling summary() in a code chunk will work but this will give you quite
-an ugly result (just try it for yourself!). When it comes to presenting
-more structured objects like summaries, model results or for example
-correlation matrices the stargazer package is well suited.
-
-\begin{table}[!htbp] \centering 
-  \caption{Summary} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
-Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
-\hline \\[-1.8ex] 
-Girth & 31 & 13.248 & 3.138 & 8.300 & 20.600 \\ 
-Height & 31 & 76.000 & 6.372 & 63 & 87 \\ 
-Volume & 31 & 30.171 & 16.438 & 10.200 & 77.000 \\ 
-\hline \\[-1.8ex] 
-\end{tabular} 
-\end{table}
-
-Assume we want to evaluate how the height and the volume of a typical
-cherry tree are related. We are estimating this using OLS to estimate a
-simple linear model.
-
-\begin{table}[!htbp] \centering 
-  \caption{Regression results} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{5pt}}lc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-2} 
-\\[-1.8ex] & Volume \\ 
-\hline \\[-1.8ex] 
- Height & 1.543$^{***}$ \\ 
-  & (0.384) \\ 
-  & \\ 
- Constant & $-$87.124$^{***}$ \\ 
-  & (29.273) \\ 
-  & \\ 
-\hline \\[-1.8ex] 
-Observations & 31 \\ 
-R$^{2}$ & 0.358 \\ 
-Adjusted R$^{2}$ & 0.336 \\ 
-Residual Std. Error & 13.397 (df = 29) \\ 
-F Statistic & 16.164$^{***}$ (df = 1; 29) \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-Now that we have our model we can visualize it.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{Dataset and regression}
-\end{figure}
-
-By the way: the (\LaTeX) command \texttt{\textbackslash{}pagebreak} can
-be used to force a page break.
-
-\pagebreak
-
-\hypertarget{citations}{%
-\subsection{Citations}\label{citations}}
-
-A \emph{bibtex} bibliography can be used for citations. The bibliography
-file used in this template is \texttt{references.bib} and you find it in
-the project directory. It is easily edited using \emph{RStudio}
-
-A simple \emph{bibtex} entry looks like this:
-
-\begin{verbatim}
-@book{
-  Hastie2013,
-  publisher = {Springer},
-  year = {2013},
-  title = {The elements of statistical learning},
-  author = {Hastie, Trevor and Tibshirani, Robert and Friedman, Jerome H.},
-}
-\end{verbatim}
-
-The first field (\texttt{Hastie2013}) is the identifier which allows you
-to cite a reference.
-
-You can cite a source in Harvard style like this: \autocite{Hastie2013}
-or \textcite{Hastie2013}.
-
-A cited source will be automatically added to the reference section at
-the end of the document.
-
-\pagebreak
-
-\hypertarget{conclusion}{%
-\section{\texorpdfstring{Conclusion
-\label{chap:conc}}{Conclusion }}\label{conclusion}}
-
-Your conclusion. Note that you may also refer to a specific chapter or
-section in the document, provided there is a label. We have anchored a
-label to the section header of the Conclusion. You may reference it as
-follow: See Chapter \ref{chap:conc}.
-
-This section allows you to reference R packages used in the analysis.
-Simply include them in \texttt{R\_packages.bib} as \emph{bibtex} entries
-and include the identifiers using
-\texttt{\textbackslash{}notecite\{...\}} as shown below.
-
-\pagebreak
-
-\addcontentsline{toc}{section}{References}
-\printbibliography[title = References]
-\cleardoublepage
-
-\begin{refsection}
-\nocite{R-base}
-\nocite{R-Studio}
-
-\printbibliography[title = Software-References]
-\addcontentsline{toc}{section}{Software-References}
-\end{refsection}
-
-\cleardoublepage
-\appendix
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-
-\newgeometry{top=2.5cm, left = 2cm, right = 2cm, bottom = 2.5cm}
-
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\hypertarget{description-of-relevant-variables}{%
-\subsection{\texorpdfstring{Description of relevant Variables
-\label{app:A}}{Description of relevant Variables }}\label{description-of-relevant-variables}}
-
-An example using the LaTeX environment \emph{longtable}:
-
-\begin{longtable}{l|l|p{6cm}}
-\caption{Description of relevant variables} \\
-\hline
-\textbf{Variable} & \textbf{Dataset} & \textbf{Description}    \\
-\hline
-\endfirsthead % Erster Kopf zu Ende
-% Definition des Tabellenkopfes auf den folgenden Seiten
-\caption{Description of relevant variables (continued)}\\
-\hline
-\textbf{Variable} & \textbf{Dataset} & \textbf{Description}    \\
-\hline
-\endhead % Zweiter Kopf ist zu Ende
-\multicolumn{3}{r}{Continued on the next page.}\\
-\endfoot
-\hline
-\multicolumn{3}{r}{End of table.} \\
-\endlastfoot
-Student.Pseudonym & all & Student’s immatriculation number \\
-\hline
-Semester & FS Data & Academic semester.
-The first four digits stand for the year, the fifth digit is either a 1 (summer semester) or a 2 (winter semester).
-Example: "20062" describes the winter semester 2006/2007, "20071" the summer semester 2007 \\
-\hline
-FS & FS, Pruefung Data & Study semester \\
-\hline
-Status & FS Data & Status as a student. "R" stands for re-enrollment, "N" for new enrollment, "E" for initial enrollment and "B" for leave of absence \\
-\hline
-Status.dpp & Pruefung Data & See variable Status \\
-\hline
-Austrittsgrund & FS, Studium Data & Reason for dropping out \\
-\hline
-SGCode & all & Code for study program. The last digit represents the examination regulations (PO). A change to the PO is thus considered as a new program in the system \\
-\hline
-Abschluss\_Bezeichnung & FS, Pruefung Data & Name of the degree that students receive upon successful completion of their studies \\
-\hline
-Fach\_Bezeichnung & FS, Pruefung Data & Name of the course of studies \\
-\hline
-PO & all & Examination regulations \\
-\hline
-PolyvalentePruefungsnummer & Pruefung Data & Polyvalent examination number \\
-\hline
-Bezeichnung & Pruefung Data & Name of the examination \\
-\hline
-Pruefungssemester & Pruefung Data & Semester in which the student was registered for the exam. Structure analogous to the variable ‘Semester’ in FS Data \\
-\hline
-Status.Pruefung & Pruefung Data & Status of the exam.
-"BE" means Passed, "ZU" means Withdrawn, "NB" means Not Passed and "PV" means Examination Existing (applications or performances are available, but the whole module is not yet completed) \\
-\hline
-Versuchszahl & Pruefung Data & Trial number \\
-\hline
-Note & Pruefung Data & Grade \\
-\hline
-VerbuchteECTSCP & Pruefung Data & Received ECTSCP \\
-\hline
-Start\_Semester & Studium Data & Semester in which the studies were started. Structure analogous to the variable ‘Semester’ in FS Data \\
-\hline
-Ende\_Semester & Studium Data & Semester in which the studies were finshed. Structure analogous to the variable ‘Semester’ in FS Data
-\label{tab_var}
-\end{longtable}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{elasticnet }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula, }\AttributeTok{data =}\NormalTok{ train,}
+       \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{),}
+       \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \FunctionTok{seq}\NormalTok{(}\AttributeTok{from=}\DecValTok{0}\NormalTok{, }\AttributeTok{to=}\DecValTok{1}\NormalTok{, }\AttributeTok{by =} \FloatTok{0.1}\NormalTok{),}
+       \AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\AttributeTok{from=}\DecValTok{0}\NormalTok{, }\AttributeTok{to=}\FloatTok{0.15}\NormalTok{, }\AttributeTok{by =} \FloatTok{0.001}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
 

### 2023-08-23 code/2_baseline_regressions.R
Use CV of 3

@@ -30,7 +30,7 @@ validation<- ds[-trainindex_filter, ]
 # validation_filter <- remaining_filter[valid_test_index_filter, ]
 # test_filter <- remaining_filter[-valid_test_index_filter, ]
 
-cat("training dim: ", nrow(train), ", test dim: ", nrow(remaining), "\n")
+cat("training dim: ", nrow(train), ", test dim: ", nrow(validation), "\n")
 
 full.col.names <- colnames(ds)
 occrecode.labels <- full.col.names[match("armed_forces", full.col.names):match("transportation", full.col.names)]
@@ -76,7 +76,7 @@ formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")
 mod_full <- train(formula,
                   data = train, 
                   method = "lm",  
-                  trControl = trainControl(method = "cv"))
+                  trControl = trainControl(method = "cv", number = 3))
 summary(mod_full)
 rmse <- caret::RMSE(pred = predict(mod_full, validation), obs = validation$realrinc)
 cat("Root Mean Squared Error (RMSE):", rmse, "\n") #25506.05
@@ -89,17 +89,17 @@ mod_wo_interactions <-train(realrinc ~ female + age + prestg10 + childs + armed_
                             graduate + highschool + juniorcollege +  + divorced + married + nevermarried + separated,
                             data = train, 
                             method = "lm",  
-                            trControl = trainControl(method = "cv"))
+                            trControl = trainControl(method = "cv", number = 3))
 summary(mod_wo_interactions)
 rmse <- caret::RMSE(pred = predict(mod_wo_interactions, validation), obs = validation$realrinc)
-cat("Root Mean Squared Error (RMSE):", rmse, "\n") #24611.53
+cat("Root Mean Squared Error (RMSE):", rmse, "\n") #25588.5
 
 #log(income) as outcome variable
 formula2 <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
 mod_loginc <- train(formula2,
                     data = train, 
                     method = "lm",  
-                    trControl = trainControl(method = "cv"))
+                    trControl = trainControl(method = "cv", number = 3))
 summary(mod_loginc)
 pred_log_inc <- exp(predict(mod_loginc, newdata = validation))
 rmse <- sqrt(mean((validation$realrinc - pred_log_inc)^2))

### 2023-08-23 code/4_LassoRidgeElasticNet.R

@@ -78,22 +78,22 @@ x_cols_dummys <- x_cols_dummys[!x_cols_dummys %in% c("armed_forces", "other_wrks
 #Lasso
 formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
 lasso <- train(formula, data = train, 
-                method = "glmnet", trControl = trainControl(method = "cv"), 
+                method = "glmnet", trControl = trainControl(method = "cv", number = 3), 
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
-caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.58
+caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.89
 
 
 #Ridge
 ridge <- train(formula, data = train,
-  method = "glmnet", trControl = trainControl(method = "cv"),
+  method = "glmnet", trControl = trainControl(method = "cv", number = 3),
   tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #25524.85
 caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
 
 #Elastic Net
 elasticnet <- train(formula, data = train,
-                    method = "glmnet", trControl = trainControl(method = "cv"),
+                    method = "glmnet", trControl = trainControl(method = "cv", number = 3),
                     tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
                     lambda = seq(from=0, to=0.15, by = 0.001)))
-caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #25241.7
+caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #25504.18
 
 

### 2023-08-23 Paper/paper.Rmd
CV of 3

@@
 ``` r
-formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+formula <- as.formula(paste("realrinc ~", 
+                            paste(x_cols_dummys, collapse = " + ")))
 mod_full <- train(formula,
@@
                   method = "lm",  
-                  trControl = trainControl(method = "cv"))
+                  trControl = trainControl(method = "cv", number = 3))
 ```
@@
 ``` r
-formula2 <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+formula2 <- as.formula(paste("log_realrinc ~", 
+                              paste(x_cols_dummys, collapse = " + ")))
 mod_loginc <- train(formula2,
@@
                     method = "lm",  
-                    trControl = trainControl(method = "cv"))
+                    trControl = trainControl(method = "cv", number = 3)
 ```
@@
 ridge <- train(formula, data = train,
-  method = "glmnet", trControl = trainControl(method = "cv"),
+  method = "glmnet", trControl = trainControl(method = "cv", number = 3,
   tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) 
@@
 To perform the lasso regression in `R` we set $\alpha$ equal to one.
-The lasso regression yields a lower RMSE than the ridge regression (25,504.58 vs.
+The lasso regression yields a lower RMSE than the ridge regression (25,504.89 vs.
 25,524.85) and even a lower RMSE than the linear regression.
@@
 The elastic net regularization allows $\alpha$ to vary between zero and one.
-This results in an RMSE of 25,241.7.
+This results in an RMSE of 25,504.18.
 

### 2023-08-23 Paper/paper.Rmd
minor changes in the paper

--- 2023-08-18 Paper/paper.Rmd
@@
 Some papers make use of classification techniques to predict wages [@Chakraborti].
-As we have numeric values of real wages and not a categeorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests.
+As we have numeric values of real wages and not a categorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests.
 We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler].
 Wages were already predicted using random forests by [@Eichinger].
-They find that usnig random forests results in better predictions than using linear models.
+They find that using random forests results in better predictions than using linear models.
 Additionally, they show that gender barely effects the wage predictions.
 
-Based on the previous literature, we use severall regression techniques to predict `realrinc`:
+Based on the previous literature, we use several regression techniques to predict `realrinc`:
 
@@
 
-In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) dataset.
+In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) data set.
 We fit the aforementioned regression models on the train data and evaluate the root mean squared error (RMSE) using the validation data.
@@
 
-# Data Preparation
+# Data Preparation & Feature Engineering
 
-Before conducting the regression analysis we prepared the data.
-The provided dataset has 54,842 observations and for 38.55% of the observations the outcome variable `realrinc` is missing.
-As we cannot predict the wage for those observations we removed them.
-This left us with a dataset containing 33,702 observations.
-Additionally, we do not know the occupation for some observations (`occ10 = 9997` ) and removed those.
-This diminishes the dataset to 33,244 observations.
+The initial data set contains 11 variables, which are described in the following:
 
-In this dataset still there were missing some values for some observations, e.g., number of children or occupation.
-To impute these values we used the `mice` function.
-We imputed age, occupational prestige and number of children with predictive mean matching.
-For the factor variables education and marital status we used polytomous logisitic regression.
-We receive six imputation datasets and use mean estimation for age, occupational prestige and number of children.
+-   **year**: survey year
+-   **realrinc**: respondent's base income (in constant 1986 USD)
+-   **age**: respondent's age in years
+-   **occ10**: respondent's occupation code
+-   **occrecode**: recode of the occupation code into one of 11 main categories
+-   **prestg10**: respondent's occupational prestige score
+-   **childs**: number of children
+-   **wrkstat**: work status of the respondent (full-time, part-time, temporarily not working, unemployed (laid off), etc.)
+-   **gender**: respondent's gender
+-   **educcat**: respondent's degree level (Less Than High School, High School, etc.)
+-   **maritalcat**: respondent's marital status (Married, Widowed, Divorced, etc.)
 
-With this dataset we create dummy variables for the factor variables. Additionally, we create interaction variables.
+## Data Preparation
 
-[Something about ANOVA test for interaction variables].
+Before modeling we conduct a data preparation process.
+The initial data set has 54,842 observations and the outcome variable `realrinc` has 38.55% of missing values.
+The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we decided to eliminate these records from the original data set.
+After this, the dataset contains 33,702 observations.
+Furthermore, it is impossible to deduce the occupation for some observations (`occ10 = 9997`) and we remove these from the data set.
+At the end, the final data set has 33,244 observations, and we use it for the modeling process.
 
-[Should we consider outlier handling? cf. Eichinger 2022]
+This data set still had some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others.
+We use the `mice` package with five variables for the imputation.
+The features *age, occupational prestige and number of children* are imputed using the pmm (`predictive mean matching`) method.
+We use the polytomous logistic regression method for the factor variables (*education and marital status*).
+The imputation process generates 6 different data sets.
+We apply the estimation of the mean and the majority vote to aggregate the data sets into a single data set.
+The mean estimation is applied for the numeric variables and majority vote for the categorical variables..
 
-In a final step we calculate the correlation matrix using Spearman's rank correlation coefficient.
-We remove all variables that have a higher than 90% correlation with each other.
-Our final dataset then contains 33,244 observations of 361 variables.
+## Feature Engineering
 
-[Something about how we define the x variables for the models]
+After the imputation process, we create the interaction variables.
+First, we estimate some interaction features using only the numerical variables.
+Then, we transform the numerical variables into categorical variables by dividing them into different ranges (e.g., age between 18 and 30, age between 31 and 50, etc.).
+We transform these new categorical variables and the initial categorical variables (*marital status, education, etc.*) into dummy variables.
+And finally, the interaction variables are estimated with the group of dummy variables mentioned above.
+The inclusion of the interaction of the numeric variables are validate using a linear model, and the results show that the p- values of the parameters for the interactions are less than 0.05.
+
+![table](includes/reg_table.png){width="317"}
+
+Furthermore, we estimate a correlation matrix using Spearman's rank correlation coefficient with some numeric variables in order to support the previous results.
+The correlation results are shown in the following figure.
+
+![matrix_corr](includes/corr_matrix.png)
+
+After the data preparation and feature engineering processes, we construct a final data set with 33,244 observations and 421 features.
+Due to the large number of interaction variables, we decided to reduce them and use only the interaction between a group of variables (*occrecode, educcat and prestg10*).
+After reduction, we develop the modeling process with 100 variables (numerical and interaction variables).
+The following figure shows the overview of the group of variables that were estimated and used in the models.
+
+![feature_process](includes/feature_process.png)
 
@@
 
-We used five different regression techniques to predict the individuals' wages using the final dataset.
+We use five different regression techniques to predict the individuals' wages using the final data set.
 We fit every model on our training data using the `train` function from the `caret` package.
@@
 ridge <- train(formula, data = train,
-  method = "glmnet", trControl = trainControl(method = "cv"),
+  method = "glmnet", trControl = trainControl(method = "cv", number = 3),
   tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) 

### 2023-08-23 code/2_baseline_regressions.R

@@ -73,6 +73,7 @@ x_cols_dummys <- x_cols_dummys[!x_cols_dummys %in% c("armed_forces", "other_wrks
 
 #OLS regressions
 formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+set.seed(123)
 mod_full <- train(formula,
                   data = train, 
                   method = "lm",  
@@ -83,6 +84,7 @@ cat("Root Mean Squared Error (RMSE):", rmse, "\n") #25506.05
 
 
 #exclude any interactions
+set.seed(123)
 mod_wo_interactions <-train(realrinc ~ female + age + prestg10 + childs + armed_forces + business_finance + construction_extraction + farming_fishing_and_forestry + 
                             installation_maintenance_and_repair + office_and_administrative_support + production + professional + sales + service + 
                              + full_time + housekeeper + part_time + retired + school + temporarily_not_working + unemployed_laid_off + bachelor + 
@@ -95,6 +97,7 @@ rmse <- caret::RMSE(pred = predict(mod_wo_interactions, validation), obs = valid
 cat("Root Mean Squared Error (RMSE):", rmse, "\n") #25588.5
 
 #log(income) as outcome variable
+set.seed(123)
 formula2 <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
 mod_loginc <- train(formula2,
                     data = train, 

### 2023-08-23 code/4_LassoRidgeElasticNet.R

@@ -77,19 +77,22 @@ x_cols_dummys <- x_cols_dummys[!x_cols_dummys %in% c("armed_forces", "other_wrks
 
 #Lasso
 formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+set.seed(123)
 lasso <- train(formula, data = train, 
                 method = "glmnet", trControl = trainControl(method = "cv", number = 3), 
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
-caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.89
+caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.39
 
 
 #Ridge
+set.seed(123)
 ridge <- train(formula, data = train,
   method = "glmnet", trControl = trainControl(method = "cv", number = 3),
   tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #25524.85
 caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
 
 #Elastic Net
+set.seed(123)
 elasticnet <- train(formula, data = train,
                     method = "glmnet", trControl = trainControl(method = "cv", number = 3),
                     tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),

### 2023-08-28 Paper/paper.Rmd
GB and RF results

@@
 Additionally, they show that gender barely effects the wage predictions.
+Likewise, the boosting methods used by [@Chakrabarty2018] in salary level classification problems showed better performance compared to a traditional algorithm such as SVM.
+Furthermore, [@BonaccoltoToepfer1612] suggests the use of a penalization approach for high dimensional problems in his research about Gender wage inequality.
 
@@
 ```
+
+## Random Forest
+
+According to the literature review, tree-based models perform considerably well in range in predicting income ranges (classification problems).
+For this reason, the Random forest algorithm is used for modeling.
+This experiment use the income without logarithm transformation, the initial features and some interaction variables.
+
+``` r
+tnr_rdgrid_search = tnr("random_search", batch_size = 10)
+rsmp_cv3 = rsmp("cv", folds = 3)
+msr_ce = msr("regr.rmse")
+learner = lrn("regr.lightgbm",
+              boosting = "rf",
+              objective = "regression",
+              max_depth = to_tune(seq(5,12,1)),
+              num_leaves = to_tune(seq(270,280,1)),
+              num_iterations  = to_tune(seq(132,137,1)),
+              min_data_in_leaf = to_tune(seq(70,85,1)),
+              min_data_in_bin = to_tune(seq(10,15, 1)),
+              feature_fraction_bynode = to_tune(seq(0.3,0.4,0.1)),
+              bagging_fraction = to_tune(seq(0.2,0.3,0.1)),
+              bagging_freq = to_tune(seq(3,5,1)),
+              feature_fraction = to_tune(seq(0.7,0.9,0.1)),
+              convert_categorical = TRUE,
+              force_col_wise = TRUE,
+              verbose = 1,
+              num_threads = 5
+)
+
+instance.rf = tune(
+  tuner = tnr_rdgrid_search,
+  task = task,
+  learner = learner,
+  resampling = rsmp_cv3,
+  measures = msr_ce,
+  term_evals = 150,
+  store_models = FALSE
+)
+```
+
+Due to the number of hyper parameters in Random Forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
+After many attempts, we found the best hyper parameters that produced the lowest error.
+The hyper parameters tuned were `max_depth`, `num_leaves`, `num_iterations` and others.
+In addition, we used a random search approach in the tuning process with a cross-validation of 3.
+Once we found the model along with the hyper parameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test data set and obtained a RMSE of 31745.54.
+The hyper parameters of the best model are shown in the following table.
+
+[best hyper parameter of RF]
+
+In addition, we can analyze the importance of the features within the modeling process using Random Forest, and these results are shown in the following table.
+
+[feature importance of RF]
+
+## Gradient Boosting
+
+Gradient Boosting also has many hyper parameters like Random Forest and we decided to apply a similar approach to the previous algorithm to find the best model with the hyper parameters (random search approach in the tuning process with a cross-validation of 3).
+Furthermore, the experiments with Gradient Boosting used the target variable without logarithm transformation.
+
+``` r
+tnr_rdgrid_search = tnr("random_search", batch_size = 10)
+rsmp_cv3 = rsmp("cv", folds = 3)
+msr_ce = msr("regr.rmse")
+learner.gb = lrn("regr.lightgbm",
+                 boosting = "gbdt",
+                 objective = "regression",
+                 max_depth = to_tune(seq(3, 6, 1)),
+                 num_leaves = to_tune(seq(6,8,1)),
+                 min_data_in_leaf = to_tune(seq(39,45,1)),
+                 min_data_in_bin = to_tune(seq(8,11,1)),
+                 feature_fraction = to_tune(seq(0.3,0.5,0.1)),
+                 feature_fraction_bynode = to_tune(seq(0.1,0.2,0.1)),
+                 learning_rate = to_tune(seq(0.03, 0.04, 0.01)),
+                 num_iterations  = to_tune(seq(39,45,1)), #40
+                 lambda_l1 = to_tune(seq(0.4, 0.5, 0.1)),
+                 lambda_l2 = to_tune(seq(0.30, 0.33, 0.01)),
+                 convert_categorical = TRUE,
+                 force_col_wise = TRUE,
+                 verbose = 1,
+                 num_threads = 5
+)
+
+instance.gb = tune(
+  tuner = tnr_rdgrid_search,
+  task = task,
+  learner = learner.gb,
+  resampling = rsmp_cv3,
+  measures = msr_ce,
+  term_evals = 170,
+  store_models = FALSE
+)
+```
+
+We tried many experiments and found the best hyper parameters that produced the lowest error.
+In Gradient Boosting modeling we adjusted the hyper parameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
+Once the best model was found, it obtained an RMSE of 29207.55 on the test data set. The hyper parameters of the best model are shown in the following table.
+
+[best hyper parameter of GB]
+
+Furthermore, the following table shows the importance of the features within the modeling process using Gradient Boosting.
+
+[feature importance of GB]
+
+These results show that in tree-based methods, the Gradient Boosting performance overcomes Random Forest, in other words, the error of Gradient Boosting is approximately 8% lower than the error of Random Forest.
+In addition, the error of these model are large than the linear models.

### 2023-08-28 Paper/paper.Rmd
using log income everywhere

@@
 geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
-biblio-files: references.bib
+biblio-files: ref.bib
 classoption: a4paper
@@
 
-[Write something about cross validation?]
+Furthermore, we use cross validation (CV) to assess the performance of a model and to select appropriate hyperparameters. CV prevents overfitting and helps in model selection.
 
@@
 -   **year**: survey year
+
 -   **realrinc**: respondent's base income (in constant 1986 USD)
+
 -   **age**: respondent's age in years
+
 -   **occ10**: respondent's occupation code
+
 -   **occrecode**: recode of the occupation code into one of 11 main categories
+
 -   **prestg10**: respondent's occupational prestige score
+
 -   **childs**: number of children
+
 -   **wrkstat**: work status of the respondent (full-time, part-time, temporarily not working, unemployed (laid off), etc.)
+
 -   **gender**: respondent's gender
+
 -   **educcat**: respondent's degree level (Less Than High School, High School, etc.)
+
 -   **maritalcat**: respondent's marital status (Married, Widowed, Divorced, etc.)
@@
 
-![table](includes/reg_table.png){width="317"}
+![Linear Model: Inclusion of Interaction Variables](includes/reg_table.png){width="317"}
 
@@
 
-![matrix_corr](includes/corr_matrix.png)
+![Correlation Matrix](includes/corr_matrix.png)
 
@@
 
-![feature_process](includes/feature_process.png)
+![Feature Engineering](includes/feature_process.png)
 
@@
 This results in a higher RMSE than the linear model (25,909.82).
+Still, the linear model yields predictions of `realrinc` that are negative, what is implausible.
+Using the log of `realrinc` only positive predictions are generated.
+Therefore, we prefer using the log for the following predictions.
 
@@
 $$
-RSS + \lambda \sum^p _{j=1} \beta^2_j 
+y = RSS + \lambda \sum^p _{j=1} \beta^2_j 
 $$
@@
   method = "glmnet", trControl = trainControl(method = "cv", number = 3),
-  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) 
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0,1,0.001)))
 ```
 
-Perform a ridge regression to predict real incomes yields an RMSE of 25,524.85.
+Perform a ridge regression to predict real incomes yields an RMSE of 26,062.58.
 
@@
        method = "glmnet", trControl = trainControl(method = "cv"), 
-       tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
+       tuneGrid = expand.grid(alpha = 1, lambda = seq(0,1,0.001)))
 ```
 
-To perform the lasso regression in `R` we set $\alpha$ equal to one.
-The lasso regression yields a lower RMSE than the ridge regression (25,504.89 vs.
-25,524.85) and even a lower RMSE than the linear regression.
+To perform the lasso regression in `R` we set $\alpha$ equal to one. The lasso regression yields a lower RMSE than the ridge regression (25,918.25 vs. 26,062.58).
 
@@
 The elastic net regularization allows $\alpha$ to vary between zero and one.
-This results in an RMSE of 25,504.18.
+This results in an RMSE of 25,918.81.
 
@@
 According to the literature review, tree-based models perform considerably well in range in predicting income ranges (classification problems).
-For this reason, the Random forest algorithm is used for modeling.
+For this reason, we use the random forest algorithm for modeling.
 This experiment use the income without logarithm transformation, the initial features and some interaction variables.
@@
 
-Due to the number of hyper parameters in Random Forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
+Due to the number of hyper parameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
 After many attempts, we found the best hyper parameters that produced the lowest error.
@@
 In addition, we used a random search approach in the tuning process with a cross-validation of 3.
-Once we found the model along with the hyper parameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test data set and obtained a RMSE of 31745.54.
+Once we found the model along with the hyper parameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test data set and obtained a RMSE of 31,745.54.
 The hyper parameters of the best model are shown in the following table.
@@
 
-In addition, we can analyze the importance of the features within the modeling process using Random Forest, and these results are shown in the following table.
+In addition, we can analyze the importance of the features within the modeling process using random forest, and these results are shown in the following table.
 
@@
 
-Gradient Boosting also has many hyper parameters like Random Forest and we decided to apply a similar approach to the previous algorithm to find the best model with the hyper parameters (random search approach in the tuning process with a cross-validation of 3).
-Furthermore, the experiments with Gradient Boosting used the target variable without logarithm transformation.
+Gradient Boosting also has many hyper parameters like random forest and we decided to apply a similar approach to the previous algorithm to find the best model with the hyper parameters (random search approach in the tuning process with a cross-validation of 3).
+Furthermore, the experiments with Gradient Boosting use the target variable without logarithm transformation.
 
@@
 In Gradient Boosting modeling we adjusted the hyper parameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
-Once the best model was found, it obtained an RMSE of 29207.55 on the test data set. The hyper parameters of the best model are shown in the following table.
+Once the best model was found, it obtained an RMSE of 29,207.55 on the test dataset.
+The hyper parameters of the best model are shown in the following table.
 
@@
 These results show that in tree-based methods, the Gradient Boosting performance overcomes Random Forest, in other words, the error of Gradient Boosting is approximately 8% lower than the error of Random Forest.
-In addition, the error of these model are large than the linear models.
+In addition, the errors of these models are larger than the linear models.

### 2023-08-28 Paper/paper.tex
using log income everywhere

@@
 \addbibresource{R_packages.bib}
-\bibliography{references.bib}
+\bibliography{ref.bib}
 \usepackage{color}
@@
 
-  Ape Monkey
+ Lea Bergmann
 
@@
 
-  Ape@Monkey.biz
+lea.bergmann@rwi-essen.de
+lea.bergmann.1j@stud.uni-due.de
 
-  M.Sc. Economics
+PhD Economics
 
-  \nth{5}
+  \nth{1}
 
-  Summer Term 2020
+  Summer Term 2023
 
@@
   \nth{4}
-
-  Summer Term 2020
-
-  \columnbreak
-
-  Darth Vader
-
-  543556
-
-  vader@emperialenterprises.com
-
-  M.Sc. Economics
-
-  \nth{3}
 
@@
 using linear models. Additionally, they show that gender barely effects
-the wage predictions.
+the wage predictions. Likewise, the boosting methods used by
+\autocite{Chakrabarty2018} in salary level classification problems
+showed better performance compared to a traditional algorithm such as
+SVM. Furthermore, \autocite{BonaccoltoToepfer1612} suggests the use of a
+penalization approach for high dimensional problems in his research
+about Gender wage inequality.
 
@@
 The initial data set contains 11 variables, which are described in the
-following: - \textbf{year}: survey year - \textbf{realrinc}:
-respondent's base income (in constant 1986 USD) - \textbf{age}:
-respondent's age in years - \textbf{occ10}: respondent's occupation code
-- \textbf{occrecode}: recode of the occupation code into one of 11 main
-categories - \textbf{prestg10}: respondent's occupational prestige score
-- \textbf{childs}: number of children - \textbf{wrkstat}: work status of
-the respondent (full-time, part-time, temporarily not working,
-unemployed (laid off), etc.) - \textbf{gender}: respondent's gender -
-\textbf{educcat}: respondent's degree level (Less Than High School, High
-School, etc.) - \textbf{maritalcat}: respondent's marital status
-(Married, Widowed, Divorced, etc.)
+following:
+
+\begin{itemize}
+\item
+  \textbf{year}: survey year
+\item
+  \textbf{realrinc}: respondent's base income (in constant 1986 USD)
+\item
+  \textbf{age}: respondent's age in years
+\item
+  \textbf{occ10}: respondent's occupation code
+\item
+  \textbf{occrecode}: recode of the occupation code into one of 11 main
+  categories
+\item
+  \textbf{prestg10}: respondent's occupational prestige score
+\item
+  \textbf{childs}: number of children
+\item
+  \textbf{wrkstat}: work status of the respondent (full-time, part-time,
+  temporarily not working, unemployed (laid off), etc.)
+\item
+  \textbf{gender}: respondent's gender
+\item
+  \textbf{educcat}: respondent's degree level (Less Than High School,
+  High School, etc.)
+\item
+  \textbf{maritalcat}: respondent's marital status (Married, Widowed,
+  Divorced, etc.)
+\end{itemize}
 
@@
 
-Before modeling we prepared the data. The initial data set has 54,842
-observations and the outcome variable \texttt{realrinc} has 38.55\% of
-missing values. The imputation of the target variable was discarded to
-avoid generating a bias in the modeling and therefore we decided to
-eliminate these records from the original data set. After this, the data
-set contains 33,702 observations. Furthermore, it was impossible to
-deduce the occupation for some observations (\texttt{occ10\ =\ 9997})
-and these are removed from the data set. At the end, the final data set
-has 33,244 observations, and is used for the modeling process.
+Before modeling we conduct a data preparation process. The initial data
+set has 54,842 observations and the outcome variable \texttt{realrinc}
+has 38.55\% of missing values. The imputation of the target variable is
+discarded to avoid generating a bias in the modeling and therefore we
+decided to eliminate these records from the original data set. After
+this, the dataset contains 33,702 observations. Furthermore, it is
+impossible to deduce the occupation for some observations
+(\texttt{occ10\ =\ 9997}) and we remove these from the data set. At the
+end, the final data set has 33,244 observations, and we use it for the
+modeling process.
 
-This data set still has some missing values for the regressors, e.g.,
-number of children, age, marital status, degree level, and others. Using
-the \texttt{mice} package we conduct the imputation using five
-variables. We impute the features \emph{age, occupational prestige and
-number of children} using the pmm (\texttt{predictive\ mean\ matching})
-method. The polytomous logistic regression method is used for the factor
+This data set still had some missing values for the regressors, e.g.,
+number of children, age, marital status, degree level, and others. We
+use the \texttt{mice} package with five variables for the imputation.
+The features \emph{age, occupational prestige and number of children}
+are imputed using the pmm (\texttt{predictive\ mean\ matching}) method.
+We use the polytomous logistic regression method for the factor
 variables (\emph{education and marital status}). The imputation process
-uses a number of multiple imputations of six. This means that the
-imputation process creates 6 different data sets, and then the
-estimation of the mean and the majority vote is applied to aggregate the
-data sets into a single data set. The mean estimation is applied for
-numeric variables and majority vote for categorical variables..
+generates 6 different data sets. We apply the estimation of the mean and
+the majority vote to aggregate the data sets into a single data set. The
+mean estimation is applied for the numeric variables and majority vote
+for the categorical variables..
 
@@
 
-After the imputation process we create interaction variables. First,
-some interaction features are estimated using only the numerical
-variables, and then we transform the numerical variables into
-categorical variables by dividing them into different ranges (e.g.~age
-between 18 and 30, age between 31 and 50, etc). These new categorical
+After the imputation process, we create the interaction variables.
+First, we estimate some interaction features using only the numerical
+variables. Then, we transform the numerical variables into categorical
+variables by dividing them into different ranges (e.g., age between 18
+and 30, age between 31 and 50, etc.). We transform these new categorical
 variables and the initial categorical variables (\emph{marital status,
-education, etc}) are transformed into dummy variables. And finally, we
-estimate the interaction variables with the group of dummy variables
-mentioned above. We validate the inclusion of the interaction of the
-numeric variables using a linear model. The results show that the
-p-values of the parameters for the interactions are less than 0.05.
+education, etc.}) into dummy variables. And finally, the interaction
+variables are estimated with the group of dummy variables mentioned
+above. The inclusion of the interaction of the numeric variables are
+validate using a linear model, and the results show that the p- values
+of the parameters for the interactions are less than 0.05.
 
@@
 \centering
-\includegraphics{includes/reg_table.png}
-\caption{table}
+\includegraphics[width=3.30208in,height=\textheight]{includes/reg_table.png}
+\caption{Linear Model: Inclusion of Interaction Variables}
 \end{figure}
@@
 \includegraphics{includes/corr_matrix.png}
-\caption{matrix\_corr}
+\caption{Correlation Matrix}
 \end{figure}
@@
 them and use only the interaction between a group of variables
-(\emph{occrecode, educcat and prestg10}). After reduction, the modeling
-process is developed with 100 variables (numerical and interaction
+(\emph{occrecode, educcat and prestg10}). After reduction, we develop
+the modeling process with 100 variables (numerical and interaction
 variables). The following figure shows the overview of the group of
-variables that we estimated and used in the models.
+variables that were estimated and used in the models.
 
@@
 \includegraphics{includes/feature_process.png}
-\caption{feature\_process}
+\caption{Feature Engineering}
 \end{figure}
@@
 We use five different regression techniques to predict the individuals'
-wages using the final dataset. We fit every model on our training data
+wages using the final data set. We fit every model on our training data
 using the \texttt{train} function from the \texttt{caret} package. To
@@
 \begin{Highlighting}[]
-\NormalTok{formula }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"realrinc \textasciitilde{}"}\NormalTok{, }\FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
+\NormalTok{formula }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"realrinc \textasciitilde{}"}\NormalTok{, }
+                            \FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
 \NormalTok{mod\_full }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula,}
@@
                   \AttributeTok{method =} \StringTok{"lm"}\NormalTok{,  }
-                  \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{))}
+                  \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{, }\AttributeTok{number =} \DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
 \texttt{realrinc} as our dependent variable. This results in a higher
-RMSE than the linear model (25,909.82).
+RMSE than the linear model (25,909.82). Still, the linear model yields
+predictions of \texttt{realrinc} that are negative, what is implausible.
+Using the log of \texttt{realrinc} only positive predictions are
+generated. Therefore, we prefer using the log for the following
+predictions.
 
@@
 \begin{Highlighting}[]
-\NormalTok{formula2 }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"log\_realrinc \textasciitilde{}"}\NormalTok{, }\FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
+\NormalTok{formula2 }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"log\_realrinc \textasciitilde{}"}\NormalTok{, }
+                              \FunctionTok{paste}\NormalTok{(x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{)))}
 \NormalTok{mod\_loginc }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula2,}
@@
                     \AttributeTok{method =} \StringTok{"lm"}\NormalTok{,  }
-                    \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{))}
+                    \AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{, }\AttributeTok{number =} \DecValTok{3}\NormalTok{)}
 \end{Highlighting}
@@
 \[
-RSS + \lambda \sum^p _{j=1} \beta^2_j 
+y = RSS + \lambda \sum^p _{j=1} \beta^2_j 
 \]
@@
 \NormalTok{ridge }\OtherTok{\textless{}{-}} \FunctionTok{train}\NormalTok{(formula, }\AttributeTok{data =}\NormalTok{ train,}
-  \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{),}
-  \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{0}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{500}\NormalTok{, }\DecValTok{1}\NormalTok{))) }
+  \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{, }\AttributeTok{number =} \DecValTok{3}\NormalTok{),}
+  \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{0}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.001}\NormalTok{)))}
 \end{Highlighting}
@@
 Perform a ridge regression to predict real incomes yields an RMSE of
-25,524.85.
+26,062.58.
 
@@
        \AttributeTok{method =} \StringTok{"glmnet"}\NormalTok{, }\AttributeTok{trControl =} \FunctionTok{trainControl}\NormalTok{(}\AttributeTok{method =} \StringTok{"cv"}\NormalTok{), }
-       \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{1}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{500}\NormalTok{,}\DecValTok{1}\NormalTok{)))}
+       \AttributeTok{tuneGrid =} \FunctionTok{expand.grid}\NormalTok{(}\AttributeTok{alpha =} \DecValTok{1}\NormalTok{, }\AttributeTok{lambda =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{,}\FloatTok{0.001}\NormalTok{)))}
 \end{Highlighting}
@@
 one. The lasso regression yields a lower RMSE than the ridge regression
-(25,504.58 vs.~25,524.85) and even a lower RMSE than the linear
-regression.
+(25,918.25 vs.~26,062.58).
 
@@
 regularization allows \(\alpha\) to vary between zero and one. This
-results in an RMSE of 25,241.7.
+results in an RMSE of 25,918.81.
 
@@
 
+\hypertarget{random-forest}{%
+\subsection{Random Forest}\label{random-forest}}
+
+According to the literature review, tree-based models perform
+considerably well in range in predicting income ranges (classification
+problems). For this reason, we use the random forest algorithm for
+modeling. This experiment use the income without logarithm
+transformation, the initial features and some interaction variables.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tnr\_rdgrid\_search }\OtherTok{=} \FunctionTok{tnr}\NormalTok{(}\StringTok{"random\_search"}\NormalTok{, }\AttributeTok{batch\_size =} \DecValTok{10}\NormalTok{)}
+\NormalTok{rsmp\_cv3 }\OtherTok{=} \FunctionTok{rsmp}\NormalTok{(}\StringTok{"cv"}\NormalTok{, }\AttributeTok{folds =} \DecValTok{3}\NormalTok{)}
+\NormalTok{msr\_ce }\OtherTok{=} \FunctionTok{msr}\NormalTok{(}\StringTok{"regr.rmse"}\NormalTok{)}
+\NormalTok{learner }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
+              \AttributeTok{boosting =} \StringTok{"rf"}\NormalTok{,}
+              \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
+              \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{270}\NormalTok{,}\DecValTok{280}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{132}\NormalTok{,}\DecValTok{137}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{70}\NormalTok{,}\DecValTok{85}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{10}\NormalTok{,}\DecValTok{15}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.4}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{bagging\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.2}\NormalTok{,}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{bagging\_freq =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.7}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
+              \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
+              \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
+              \AttributeTok{num\_threads =} \DecValTok{5}
+\NormalTok{)}
+
+\NormalTok{instance.rf }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
+  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
+  \AttributeTok{task =}\NormalTok{ task,}
+  \AttributeTok{learner =}\NormalTok{ learner,}
+  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
+  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
+  \AttributeTok{term\_evals =} \DecValTok{150}\NormalTok{,}
+  \AttributeTok{store\_models =} \ConstantTok{FALSE}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Due to the number of hyper parameters in random forest, we use the
+\texttt{lightgbm} and \texttt{mlr3} packages to model and tune the hyper
+parameter in a time-optimal way. After many attempts, we found the best
+hyper parameters that produced the lowest error. The hyper parameters
+tuned were \texttt{max\_depth}, \texttt{num\_leaves},
+\texttt{num\_iterations} and others. In addition, we used a random
+search approach in the tuning process with a cross-validation of 3. Once
+we found the model along with the hyper parameters with lowest error in
+the cross-validation, we proceeded to evaluate this model on the test
+data set and obtained a RMSE of 31,745.54. The hyper parameters of the
+best model are shown in the following table.
+
+{[}best hyper parameter of RF{]}
+
+In addition, we can analyze the importance of the features within the
+modeling process using random forest, and these results are shown in the
+following table.
+
+{[}feature importance of RF{]}
+
+\hypertarget{gradient-boosting}{%
+\subsection{Gradient Boosting}\label{gradient-boosting}}
+
+Gradient Boosting also has many hyper parameters like random forest and
+we decided to apply a similar approach to the previous algorithm to find
+the best model with the hyper parameters (random search approach in the
+tuning process with a cross-validation of 3). Furthermore, the
+experiments with Gradient Boosting use the target variable without
+logarithm transformation.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tnr\_rdgrid\_search }\OtherTok{=} \FunctionTok{tnr}\NormalTok{(}\StringTok{"random\_search"}\NormalTok{, }\AttributeTok{batch\_size =} \DecValTok{10}\NormalTok{)}
+\NormalTok{rsmp\_cv3 }\OtherTok{=} \FunctionTok{rsmp}\NormalTok{(}\StringTok{"cv"}\NormalTok{, }\AttributeTok{folds =} \DecValTok{3}\NormalTok{)}
+\NormalTok{msr\_ce }\OtherTok{=} \FunctionTok{msr}\NormalTok{(}\StringTok{"regr.rmse"}\NormalTok{)}
+\NormalTok{learner.gb }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
+                 \AttributeTok{boosting =} \StringTok{"gbdt"}\NormalTok{,}
+                 \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
+                 \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{6}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{8}\NormalTok{,}\DecValTok{11}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+                 \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}\FloatTok{0.2}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+                 \AttributeTok{learning\_rate =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.03}\NormalTok{, }\FloatTok{0.04}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
+                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)), }\CommentTok{\#40}
+                 \AttributeTok{lambda\_l1 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
+                 \AttributeTok{lambda\_l2 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.30}\NormalTok{, }\FloatTok{0.33}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
+                 \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
+                 \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
+                 \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
+                 \AttributeTok{num\_threads =} \DecValTok{5}
+\NormalTok{)}
+
+\NormalTok{instance.gb }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
+  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
+  \AttributeTok{task =}\NormalTok{ task,}
+  \AttributeTok{learner =}\NormalTok{ learner.gb,}
+  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
+  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
+  \AttributeTok{term\_evals =} \DecValTok{170}\NormalTok{,}
+  \AttributeTok{store\_models =} \ConstantTok{FALSE}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+We tried many experiments and found the best hyper parameters that
+produced the lowest error. In Gradient Boosting modeling we adjusted the
+hyper parameters such as \texttt{max\_depth}, \texttt{num\_leaves},
+\texttt{min\_data\_in\_leaf}, \texttt{feature\_fraction} and others.
+Once the best model was found, it obtained an RMSE of 29,207.55 on the
+test dataset. The hyper parameters of the best model are shown in the
+following table.
+
+{[}best hyper parameter of GB{]}
+
+Furthermore, the following table shows the importance of the features
+within the modeling process using Gradient Boosting.
+
+{[}feature importance of GB{]}
+
+These results show that in tree-based methods, the Gradient Boosting
+performance overcomes Random Forest, in other words, the error of
+Gradient Boosting is approximately 8\% lower than the error of Random
+Forest. In addition, the errors of these models are larger than the
+linear models.
+
 \end{document}

### 2023-08-28 code/2_baseline_regressions.R

@@ -92,6 +92,7 @@ mod_wo_interactions <-train(realrinc ~ female + age + prestg10 + childs + armed_
                             data = train, 
                             method = "lm",  
                             trControl = trainControl(method = "cv", number = 3))
+
 summary(mod_wo_interactions)
 rmse <- caret::RMSE(pred = predict(mod_wo_interactions, validation), obs = validation$realrinc)
 cat("Root Mean Squared Error (RMSE):", rmse, "\n") #25588.5

### 2023-08-28 code/4_LassoRidgeElasticNet.R

@@ -76,20 +76,20 @@ x_cols_dummys <- x_cols_dummys[!x_cols_dummys %in% c("armed_forces", "other_wrks
 
 
 #Lasso
-formula <- as.formula(paste("realrinc ~", paste(x_cols_dummys, collapse = " + ")))
+formula <- as.formula(paste("log_realrinc ~", paste(x_cols_dummys, collapse = " + ")))
 set.seed(123)
 lasso <- train(formula, data = train, 
                 method = "glmnet", trControl = trainControl(method = "cv", number = 3), 
-                tuneGrid = expand.grid(alpha = 1, lambda = seq(0,500,1)))
-caret::RMSE(pred = predict(lasso, validation), obs = validation$realrinc) #25504.39
+                tuneGrid = expand.grid(alpha = 1, lambda = seq(0,1,0.001)))
+caret::RMSE(pred = exp(predict(lasso, validation)), obs = validation$realrinc) #25918.25
 
 
 #Ridge
 set.seed(123)
 ridge <- train(formula, data = train,
   method = "glmnet", trControl = trainControl(method = "cv", number = 3),
-  tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 500, 1))) #25524.85
-caret::RMSE(pred = predict(ridge, validation), obs = validation$realrinc)
+  tuneGrid = expand.grid(alpha = 0, lambda = seq(0,1,0.001)))  #26062.58
+caret::RMSE(pred = exp(predict(ridge, validation)), obs = validation$realrinc)
 
 #Elastic Net
 set.seed(123)
@@ -97,6 +97,6 @@ elasticnet <- train(formula, data = train,
                     method = "glmnet", trControl = trainControl(method = "cv", number = 3),
                     tuneGrid = expand.grid(alpha = seq(from=0, to=1, by = 0.1),
                     lambda = seq(from=0, to=0.15, by = 0.001)))
-caret::RMSE(pred = predict(elasticnet, validation), obs = validation$realrinc) #25504.18
+caret::RMSE(pred = exp(predict(elasticnet, validation)), obs = validation$realrinc) #25918.81
 
 

### 2023-08-30 Paper/paper.Rmd
Paper: Evaluation added

@@
 
-Perform a ridge regression to predict real incomes yields an RMSE of 26,062.58.
+Performing a ridge regression to predict real incomes yields an RMSE of 26,062.58.
 
@@
 In addition, the errors of these models are larger than the linear models.
+
+# Evaluation
+
+The aim of this paper is to predict individuals' wages. In order to evaluate the best prediction method we use the RMSE. The following table demonstrates the RMSE for each used method: \medskip
+
+\begin{tabular}{l c} 
+\textbf{Method} & \textbf{RMSE} \\ 
+Linear Regression & 25,506.05 \\ 
+Logistic Regression & 25,909.82 \\ 
+Ridge & 26,062.58 \\ 
+Lasso & 25,918.25 \\
+Elastic Net & 25,918.81 \\ 
+Random Forest & 31,745.54 \\ 
+Gradient Boosting & 29,207.55 \\ 
+\end{tabular}
+
+\medskip
+Conclusively, the linear regression yields the lowest RMSE and thus appears to be the best prediction method.
+Still, it delivers negative predictions of the wage.
+Therefore, we prefer the logistic regression as the best prediction method.

### 2023-08-30 Paper/paper.tex
Paper: Evaluation added

@@
 
-{[}Write something about cross validation?{]}
+Furthermore, we use cross validation (CV) to assess the performance of a
+model and to select appropriate hyperparameters. CV prevents overfitting
+and helps in model selection.
 
@@
 
-Perform a ridge regression to predict real incomes yields an RMSE of
+Performing a ridge regression to predict real incomes yields an RMSE of
 26,062.58.
@@
 
+\hypertarget{evaluation}{%
+\section{Evaluation}\label{evaluation}}
+
+The aim of this paper is to predict individuals' wages. In order to
+evaluate the best prediction method we use the RMSE. The following table
+demonstrates the RMSE for each used method: \medskip
+
+\begin{tabular}{l c} 
+\textbf{Method} & \textbf{RMSE} \\ 
+Linear Regression & 25,506.05 \\ 
+Logistic Regression & 25,909.82 \\ 
+Ridge & 26,062.58 \\ 
+Lasso & 25,918.25 \\
+Elastic Net & 25,918.81 \\ 
+Random Forest & 31,745.54 \\ 
+Gradient Boosting & 29,207.55 \\ 
+\end{tabular}
+
+\medskip
+
+Conclusively, the linear regression yields the lowest RMSE and thus
+appears to be the best prediction method. Still, it delivers negative
+predictions of the wage. Therefore, we prefer the logistic regression as
+the best prediction method.
+
 \end{document}

### 2023-08-30 Paper/paper.Rmd
update paper and code files

@@
 
+-   Generalized Additive Model
+
 In order to evaluate which of the regression techniques predicts `realrinc` the best we split the data set into a train (70%) and a test (30%) data set.
@@
 The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we decided to eliminate these records from the original data set.
-After this, the dataset contains 33,702 observations.
+After this, the data set contains 33,702 observations.
 Furthermore, it is impossible to deduce the occupation for some observations (`occ10 = 9997`) and we remove these from the data set.
-At the end, the final data set has 33,244 observations, and we use it for the modeling process.
+Finally, the data set has 33,244 observations and we use it for the modeling process.
 
 This data set still had some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others.
-We use the `mice` package with five variables for the imputation.
-The features *age, occupational prestige and number of children* are imputed using the pmm (`predictive mean matching`) method.
-We use the polytomous logistic regression method for the factor variables (*education and marital status*).
-The imputation process generates 6 different data sets.
-We apply the estimation of the mean and the majority vote to aggregate the data sets into a single data set.
-The mean estimation is applied for the numeric variables and majority vote for the categorical variables..
+So, we use the `mice` package with five variables for the imputation process.
+The features *age, occupational prestige and number of children* are imputed using the predictive mean matching (pmm) method, and we use the polytomous logistic regression method for the factor variables (*education and marital status*).
+Due to we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets.
+This parameter refers to the number of imputed data sets.
+After this, we apply the calculation of the mean and the majority vote to aggregate the data sets into one.
+Mean estimation is applied for numerical variables and majority voting for categorical variables
 
@@
 
-After the imputation process, we create the interaction variables.
-First, we estimate some interaction features using only the numerical variables.
-Then, we transform the numerical variables into categorical variables by dividing them into different ranges (e.g., age between 18 and 30, age between 31 and 50, etc.).
-We transform these new categorical variables and the initial categorical variables (*marital status, education, etc.*) into dummy variables.
-And finally, the interaction variables are estimated with the group of dummy variables mentioned above.
-The inclusion of the interaction of the numeric variables are validate using a linear model, and the results show that the p- values of the parameters for the interactions are less than 0.05.
+After the imputation process, we created some interaction variables.
+First, we estimated some interaction features using only the numerical variables.
+Then, we transformed the numerical variables (*age, child and prestg*) into categorical variables by separating them into ranges.
+The ranges created of these variables were: *age between 18 and 30, age between 31 and 50, and age greater than 50, child between 0 and 2, child between 3 and 5, child greater than 5; and prestg10 between 16 and 30, prestg10 between 31 and 50, prestg10 greater than 50*.
+After this, we transform these new categorical variables and the initial categorical variables (*marital status, education, etc.*) into dummy variables.
+Finally, the interaction variables are estimated with the group of dummy variables mentioned above (the interaction consists of the multiplication between 2 dummy variables).
+The inclusion of the interaction of the numeric variables are validate using a linear model, and the results show that the p-values of the parameters for the interactions are less than 0.05.
 
@@
 Furthermore, we estimate a correlation matrix using Spearman's rank correlation coefficient with some numeric variables in order to support the previous results.
-The correlation results are shown in the following figure.
+The correlation results are exposed in the figure below, and show that *prestg10*, *age* and *interaction between age and prest10* have the highest positive correlation with *realinc*, meanwhile the dummy variable *female* has a moderate negative correlation with the target variable.
 
@@
 
-After the data preparation and feature engineering processes, we construct a final data set with 33,244 observations and 421 features.
-Due to the large number of interaction variables, we decided to reduce them and use only the interaction between a group of variables (*occrecode, educcat and prestg10*).
-After reduction, we develop the modeling process with 100 variables (numerical and interaction variables).
+Due to the correlation assesses the linear relationship between two variables and it only uses numerical features, we decided to estimate the Predictive Power Score (PPS).
+It focuses on these limitations and allows us to have a predictive power of a regressor on a target variable.
+This measure is a normalized metric and its values are between 0 (no predictive power) and 1 (perfect predictive power).
+The PPS is calculated by comparing the performance of a model that uses a regressor `x` to predict target variable `y` versus the performance of a base model (naive).
+When the target variable is a numerical feature it is common to use a Decision Tree Regressor to predict `y` based on `x`, while in categorical cases a Decision Tree Classifier is used.
+On the other hand, base models are often defined as mean prediction or random or modal prediction for regression and classification problems, respectively.
+The PPS results are presented in the figure below, and show that *prestg10*, *wrkstat*, *educcat* and *occrecode* have the highest predictive power over *realinc* relative to the other features.
+These results help and provide some insights in the data exploration process, but we must beware of the limitations of this approach.
+
+![PPS Results](includes/pps.png)
+
+After the data preparation and feature engineering processes, we have a final data set with 33,244 observations and 421 features.
+Due to the large number of interaction variables, we decided to reduce them and use only the interaction between a group of variables (*occrecode*, *educcat* and categorical variable of *prestg10*).
+After this, we develop the modeling process with 100 variables (numerical and interaction variables).
 The following figure shows the overview of the group of variables that were estimated and used in the models.
@@
 
-According to the literature review, tree-based models perform considerably well in range in predicting income ranges (classification problems).
+According to the literature review, tree-based models perform considerably well in predicting income ranges (classification problems) [@Chakrabarty2018].
 For this reason, we use the random forest algorithm for modeling.
-This experiment use the income without logarithm transformation, the initial features and some interaction variables.
+In this experiment, income without logarithm transformation, initial variables and some interaction variables are used.
+Due to the amount of hyperparameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
+After many attempts, we found the best hyper parameters that produced the lowest error.
+The hyperparameters tuned were `max_depth`, `num_leaves`, `num_iterations`, etc, and the search space of them is showed below.
+In addition, we used a random search approach in the tuning process with a cross-validation of 3 folds.
 
@@
   measures = msr_ce,
-  term_evals = 150,
+  term_evals = 300,
   store_models = FALSE
@@
 
-Due to the number of hyper parameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
-After many attempts, we found the best hyper parameters that produced the lowest error.
-The hyper parameters tuned were `max_depth`, `num_leaves`, `num_iterations` and others.
-In addition, we used a random search approach in the tuning process with a cross-validation of 3.
-Once we found the model along with the hyper parameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test data set and obtained a RMSE of 31,745.54.
-The hyper parameters of the best model are shown in the following table.
+Once we found the model along with the hyperparameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test dataset and obtained a RMSE of 31,745.54.
+The hyperparameters of the best model are shown in the following figure.
 
-[best hyper parameter of RF]
+![Best hyperparameters for Random Forest](includes/RF_hyper.png)
 
-In addition, we can analyze the importance of the features within the modeling process using random forest, and these results are shown in the following table.
+In addition, we can analyze the features importance within the modeling process using random forest, and these results are exposed in the figure below.
 
-[feature importance of RF]
+![Top 20 of Features Importance for Random Forest](includes/RF_importance.png)
+
+Unlike the linear models presented previously, random forest does not predict negative values when logarithmic transformation is not used.
+We carry out an experiment with a random forest but using the logarithmic transformation in the target variable and keeping the same configuration (CV of 3 folds and random search approach in the tuning process).
+
+```r
+learner.wlog = lrn("regr.lightgbm",
+              boosting = "rf",
+              objective = "regression",
+              max_depth = to_tune(seq(7,12,1)),
+              num_leaves = to_tune(seq(120,140,1)),
+              max_bin = to_tune(seq(45,60,1)),
+              num_iterations  = to_tune(seq(50,90,1)),
+              min_data_in_leaf = to_tune(seq(20,35,1)), 
+              min_data_in_bin = to_tune(seq(35,50, 1)),
+              feature_fraction_bynode = to_tune(seq(0.4,0.7,0.1)),
+              bagging_fraction = to_tune(seq(0.6,0.9,0.1)),
+              bagging_freq = to_tune(seq(3,12,1)),
+              feature_fraction = to_tune(seq(0.8,0.9,0.1)),
+              convert_categorical = TRUE,
+              force_col_wise = TRUE,
+              verbose = 1,
+              num_threads = 5,
+              seed = 123
+)
+
+```
+
+The best model using the logarithmic transformation had different hyperparameters relative to the previous experiment and obtained an RMSE of 34,293.33 on the test dataset.
+This result shows that the error of this model is greater than the best model without the transformation (31,745.54).
 
@@
 
-Gradient Boosting also has many hyper parameters like random forest and we decided to apply a similar approach to the previous algorithm to find the best model with the hyper parameters (random search approach in the tuning process with a cross-validation of 3).
-Furthermore, the experiments with Gradient Boosting use the target variable without logarithm transformation.
+Gradient Boosting also has a similar amount of hyperparameters as random forest and we decided to keep a similar configuration as the previous algorithm to find the best model with the hyperparameters (random search approach in the tuning process with 3-fold cross-validation).
+This first experiment with Gradient Boosting uses the target variable without logarithmic transformation.
 
@@
                  learning_rate = to_tune(seq(0.03, 0.04, 0.01)),
-                 num_iterations  = to_tune(seq(39,45,1)), #40
+                 num_iterations  = to_tune(seq(39,45,1)),
                  lambda_l1 = to_tune(seq(0.4, 0.5, 0.1)),
@@
   measures = msr_ce,
-  term_evals = 170,
+  term_evals = 300,
   store_models = FALSE
@@
 
-We tried many experiments and found the best hyper parameters that produced the lowest error.
-In Gradient Boosting modeling we adjusted the hyper parameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
-Once the best model was found, it obtained an RMSE of 29,207.55 on the test dataset.
-The hyper parameters of the best model are shown in the following table.
+We tried many combinations of hyperparameters and found the best model that produced the least error.
+In Gradient Boosting we adjusted some hyperparameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
+Once the best model was found, it was evaluated on the test data set and an RMSE of 29,207.55 was obtained.
+The hyperparameters of the best model are shown in the following table.
 
-[best hyper parameter of GB]
+![Best hyperparameters for Gradient Boosting](includes/GB_hyper.png)
 
-Furthermore, the following table shows the importance of the features within the modeling process using Gradient Boosting.
+Gradient Boosting also provides information on the importance of features during the modeling process in this experiment.
+The following figure shows the 20 most important variables according to the Gini metric.
 
-[feature importance of GB]
+![Top 20 of Features Importance for Gradient Boosting](includes/GB_importance.png)
 
-These results show that in tree-based methods, the Gradient Boosting performance overcomes Random Forest, in other words, the error of Gradient Boosting is approximately 8% lower than the error of Random Forest.
-In addition, the errors of these models are larger than the linear models.
+Furthermore, we carried out an additional experiment with gradient boosting using the logarithm transformation on the target variable.
+
+```r
+learner.gb.wlog = lrn("regr.lightgbm",
+                 boosting = "gbdt",
+                 objective = "regression",
+                 max_depth = to_tune(seq(5, 12, 1)),
+                 max_bin = to_tune(seq(35,60,1)), 
+                 num_leaves = to_tune(seq(19,30,1)),
+                 min_data_in_leaf = to_tune(seq(30,50,1)),
+                 min_data_in_bin = to_tune(seq(20,45,1)),
+                 feature_fraction = to_tune(seq(0.3,0.8,0.1)),
+                 feature_fraction_bynode = to_tune(seq(0.7,0.9,0.1)),
+                 learning_rate = to_tune(seq(0.01, 0.1, 0.01)), 
+                 num_iterations  = to_tune(seq(60,80,1)), 
+                 lambda_l1 = to_tune(seq(0.01, 0.1, 0.01)),
+                 lambda_l2 = to_tune(seq(0.3, 0.45, 0.01)),
+                 convert_categorical = TRUE,
+                 force_col_wise = TRUE,
+                 verbose = 1,
+                 num_threads = 5,
+                 seed = 123
+)
+
+```
+
+The best model using the logarithmic transformation had a RMSE of 34,385.47 on the test dataset.
+This result shows that the error of this model is greater relative to the best model without the transformation (29,207.55).
+These results show that in tree-based methods, Gradient Boosting has a dominant performance over Random Forest.
+In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used, but the errors of these models are larger than those of the linear models.
+
+## Generalized Additive Model
+
+Generalized additive model (GAM) could be considered the next step of the linear model because it allows combining nonlinear functional forms of the features along with the linear part.
+This approach is based on relating the nonlinear predictors with a link function $g(.)$ of the expected values $E(y)$, and the following formulas offer more details about this approach.
+
+$$
+
+y \sim \text{ExpoFam}(\mu, \text{etc})\\
+E(y) = \mu \\
+g(\mu) = b_0 + f(x_1) + f(x_2) + \dots + f(x_p)
+
+$$
+
+The main difference with Generalized Lineal Models (GLM) is that this approach allows linear features to aggregate smooth functions $f(.)$ of the required variables and thus establish nonlinear relationships between regressors and the dependent variable.
+This models uses the `gam` and `mlr3` packages to model for the modeling.
+The adjusted hyperparameter was `gamma` and is responsible for producing a smoother term, and a additional penalization on the smooth term was aggregated with the parameter `select`.
+In addition, we used a 3-fold cross-validation in the training stage for this algorithm.
+This experiment uses the logarithmic transformation of the target variable, the interaction variables, smooth terms for the features *age* and *prestg10* with cubic regression splines `cr` and a value of 9 for the knots `k` (number of sections into which a feature is divided), and the smooth tensor product for the variables *age*, *prestg10* and *child* uses a thin-plate regression spline `tp` and a knot value of 1.
+The last smooth term is considered as the interaction between two features to predict the target variables.
+
+```r
+form.str.gam <- paste("log_realrinc ~ s(age, bs = 'cr', k = 9) + s(prestg10, bs = 'cr', k=9) + 
+                        ti(age,prestg10) + ti(age,childs) +",
+                       paste( x_cols_dummys, collapse = " + ") )
+form.format.gam <- as.formula(form.str.gam)
+learner.gam.wlog.smooth.tensor = lrn("regr.gam",
+                                     family = 'gaussian',
+                                     select = TRUE,
+                                     gamma = to_tune(seq(1,3,0.5)),
+                                     formula = form.format.gam
+)
+```
+
+The best model obtained a RMSE of 25,856.76 on the test dataset.
+This result shows that the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82).
+In other words, it is 0.20% better than the initial linear model.
+The figure below exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.
+
+![Approximate significance of smooth terms](includes/aprox_smooth.png)
 
@@
 Gradient Boosting & 29,207.55 \\ 
+Generalized Additive Model & 25,856.76 \\
 \end{tabular}

### 2023-08-30 code/5_models_GB_RF.R

@@ -1,5 +1,10 @@
 # set wd
-setwd("/home/user/Downloads/info_tud/statistical_learning_UDE/gender_pay_gap_project")
+tryCatch({
+  setwd("C:/Users/lbergmann/OneDrive - RWI–Leibniz-Institut für Wirtschaftsforschung e.V/Dokumente/Promotion/StatisticalLearning/gender_pay_gap_project_1")
+}, error = function(err1) {
+  # If an error occurs, handle it by setting the working directory to the second path
+  setwd("/home/user/Downloads/info_tud/statistical_learning_UDE/gender_pay_gap_project")
+})
 # import functions file
 source('code/utils_functions.R')
 
@@ -14,6 +19,7 @@ library(mlr3)
 library(mlr3learners)
 library(mlr3tuning)
 library(data.table)
+library(stargazer)
 # remotes::install_github("mlr-org/mlr3extralearners@*release")
 library(mlr3extralearners)
 
@@ -31,19 +37,6 @@ set.seed(123)
 trainindex_filter <- createDataPartition(y = ds$realrinc, p = 0.7, list = FALSE) # Split the data into training (70%) and remaining (30%)
 train_filter <- ds[trainindex_filter, ]
 remaining_filter <- ds[-trainindex_filter, ]
-
-# write.csv(rbind(train_filter %>% 
-#   select(V1) %>% 
-#   mutate(label_ds = "TRAIN"),
-# remaining_filter %>% 
-#   select(V1) %>% 
-#   mutate(label_ds = "TEST")) %>% 
-#   arrange(V1,desc(TRUE)), "data/label_dataset.csv", row.names = FALSE)
-
-
-# valid_test_index_filter <- createDataPartition(y = remaining_filter$realrinc, p = 0.5, list = FALSE) # Split the remaining data into validation (50%) and test (50%)
-# validation_filter <- remaining_filter[valid_test_index_filter, ]
-# test_filter <- remaining_filter[-valid_test_index_filter, ]
 
 cat("training dim: ", nrow(train_filter), ", test dim: ", nrow(remaining_filter), "\n")
 
@@ -66,15 +59,16 @@ for (name.col in full.col.names[c(60:ncol(ds))]){
   }
 }
 x_cols_inter <- unlist(x_cols_inter)
-  
-# create data sets with lbg format
-# select regressors and outcome
-# x_cols <- colnames(ds)[match("age", colnames(ds)):match("age_sqr", colnames(ds))]
 
[…]

### 2023-08-31 Paper/paper.Rmd
proofread papper

@@
 subtitle: "Statistical Learning"
-deadline: "tomorrow"
+deadline: "31.08.2023"
 type: "Term Paper"
-date: "today"
+date: "31.08.2023"
 supervisors: "Prof. Dr. Thomas Deckers, Karolina Gliszczynska"
@@
 The prediction of individuals' wages has been an issue in previous literature.
-As wages are relevant to companies they want to know how predict wages [@Chakraborti].
+As wages are relevant to companies they want to know how to predict wages [@Chakraborti].
 Some papers make use of classification techniques to predict wages [@Chakraborti].
@@
 Additionally, they show that gender barely effects the wage predictions.
-Likewise, the boosting methods used by [@Chakrabarty2018] in salary level classification problems showed better performance compared to a traditional algorithm such as SVM.
-Furthermore, [@BonaccoltoToepfer1612] suggests the use of a penalization approach for high dimensional problems in his research about Gender wage inequality.
+Likewise, the boosting methods used by [@Chakrabarty2018] in salary level classification problems showed better performance compared to a traditional algorithm such as support vector machines (SVM).
+Furthermore, @BonaccoltoToepfer1612 suggest the use of a penalization approach for high dimensional problems in his research about gender wage inequality.
 
@@
 
+-   Random Forest
+
 -   Gradient Boosting
-
--   Random Forest
 
@@
 The initial data set has 54,842 observations and the outcome variable `realrinc` has 38.55% of missing values.
-The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we decided to eliminate these records from the original data set.
+The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we eliminate these records from the original data set.
 After this, the data set contains 33,702 observations.
@@
 
-This data set still had some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others.
-So, we use the `mice` package with five variables for the imputation process.
+This data set still has some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others.
+Thus, we use the `mice` package with five variables for the imputation process.
 The features *age, occupational prestige and number of children* are imputed using the predictive mean matching (pmm) method, and we use the polytomous logistic regression method for the factor variables (*education and marital status*).
-Due to we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets.
+As we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets.
 This parameter refers to the number of imputed data sets.
@@
 
-After the imputation process, we created some interaction variables.
-First, we estimated some interaction features using only the numerical variables.
-Then, we transformed the numerical variables (*age, child and prestg*) into categorical variables by separating them into ranges.
-The ranges created of these variables were: *age between 18 and 30, age between 31 and 50, and age greater than 50, child between 0 and 2, child between 3 and 5, child greater than 5; and prestg10 between 16 and 30, prestg10 between 31 and 50, prestg10 greater than 50*.
-After this, we transform these new categorical variables and the initial categorical variables (*marital status, education, etc.*) into dummy variables.
-Finally, the interaction variables are estimated with the group of dummy variables mentioned above (the interaction consists of the multiplication between 2 dummy variables).
-The inclusion of the interaction of the numeric variables are validate using a linear model, and the results show that the p-values of the parameters for the interactions are less than 0.05.
+After the imputation process, we create interaction variables.
+First, we estimate some interaction features using only the numerical variables.
+Then, we transform the numerical variables (`age, childs and prestg10`) into categorical variables by separating them into ranges.
+The ranges created of these variables were: `age between 18 and 30, age between 31 and 50`, `age greater than 50; childs between 0 and 2, childs between 3 and 5, childs more than 5; prestg10 between 16 and 30, prestg10 between 31 and 50, prestg10 greater than 50`.
+After this, we transform these new categorical variables and the initial categorical variables (`marital status`, `education`, etc.) into dummy variables.
+Finally, we estimate the interaction variables with the group of dummy variables mentioned above (the interaction consists of the multiplication between two dummy variables).
+The inclusion of the interactions of the numeric variables is validate using a linear model, and the results show that the p-values of the parameters for the interactions are below 0.05.
 
@@
 Furthermore, we estimate a correlation matrix using Spearman's rank correlation coefficient with some numeric variables in order to support the previous results.
-The correlation results are exposed in the figure below, and show that *prestg10*, *age* and *interaction between age and prest10* have the highest positive correlation with *realinc*, meanwhile the dummy variable *female* has a moderate negative correlation with the target variable.
+The correlation results are exposed in figure 2, and show that `prestg10`, `age` and `interaction between age and prest10` have the highest positive correlation with `realrinc`, meanwhile the dummy variable `female` has a moderate negative correlation with the target variable.
 
@@
 
-Due to the correlation assesses the linear relationship between two variables and it only uses numerical features, we decided to estimate the Predictive Power Score (PPS).
-It focuses on these limitations and allows us to have a predictive power of a regressor on a target variable.
+As the correlation assesses the linear relationship between two variables and it only uses numerical features, we additionally estimate the Predictive Power Score (PPS).
+It overcomes these limitations and allows us to assess the predictive power of a regressor on a target variable.
 This measure is a normalized metric and its values are between 0 (no predictive power) and 1 (perfect predictive power).
@@
 On the other hand, base models are often defined as mean prediction or random or modal prediction for regression and classification problems, respectively.
-The PPS results are presented in the figure below, and show that *prestg10*, *wrkstat*, *educcat* and *occrecode* have the highest predictive power over *realinc* relative to the other features.
-These results help and provide some insights in the data exploration process, but we must beware of the limitations of this approach.
+The PPS results are presented in figure 3, and show that `prestg10`, `wrkstat`, `educcat` and `occrecode` have the highest predictive power of `realrinc` relative to the other features.
+These results provide some insights in the data exploration process, but we must beware of the limitations of this approach.
 
@@
 After the data preparation and feature engineering processes, we have a final data set with 33,244 observations and 421 features.
-Due to the large number of interaction variables, we decided to reduce them and use only the interaction between a group of variables (*occrecode*, *educcat* and categorical variable of *prestg10*).
+Due to the large number of interaction variables, we reduce them and use only the interaction between a group of variables (`occrecode`, `educcat` and categorical variable of `prestg10`).
 After this, we develop the modeling process with 100 variables (numerical and interaction variables).
-The following figure shows the overview of the group of variables that were estimated and used in the models.
+Figure 4 shows the overview of the group of estimated variables that are used in the models.
 
@@
 
-As the dataset contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection.
+As the data set contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection.
 We use lasso, ridge and elastic net regressions.
@@
 Hence, the coefficient estimates in a ridge regression minimize the sum of the residual sum of squares (RSS) and the sum of squared coefficients multiplied with a tuning parameter $\lambda$.
-We estimate $\lambda$ using cross-validation.
+We estimate $\lambda$ using CV.
 
@@
 For this reason, we use the random forest algorithm for modeling.
-In this experiment, income without logarithm transformation, initial variables and some interaction variables are used.
-Due to the amount of hyperparameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way.
-After many attempts, we found the best hyper parameters that produced the lowest error.
-The hyperparameters tuned were `max_depth`, `num_leaves`, `num_iterations`, etc, and the search space of them is showed below.
-In addition, we used a random search approach in the tuning process with a cross-validation of 3 folds.
+For this method we use income without logarithm transformation, initial variables and some interaction variables.
+Due to the amount of hyperparameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyperparameter in a time-optimal way.
+We tune the hyperparameters `max_depth`, `num_leaves`, `num_iterations`, etc.
+to find the best hyperparameters that produce the lowest error.
+The search space of the hyperparameters is showed below.
+In addition, we use a random search approach in the tuning process with a CV of 3 folds.
 
@@
 
-Once we found the model along with the hyperparameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test dataset and obtained a RMSE of 31,745.54.
-The hyperparameters of the best model are shown in the following figure.
+After identifying the model along with the hyperparameters with lowest error in CV, we proceed to evaluate this model on the test data set and obtain an RMSE of 31,745.54.
+The hyperparameters of the best model are shown in figure 5.
 
-![Best hyperparameters for Random Forest](includes/RF_hyper.png)
+```{r, echo=FALSE, out.width="50%", fig.cap="Best Hyperparameters for Random Forest"}
+knitr::include_graphics("includes/RF_hyper.png")
+```
 
-In addition, we can analyze the features importance within the modeling process using random forest, and these results are exposed in the figure below.
+In addition, we analyze the features importance within the modeling process using random forest. These results are exposed in the figure 6.
 
@@
 Unlike the linear models presented previously, random forest does not predict negative values when logarithmic transformation is not used.
-We carry out an experiment with a random forest but using the logarithmic transformation in the target variable and keeping the same configuration (CV of 3 folds and random search approach in the tuning process).
+Still, we additionally use a random forest with the logarithmic transformation in the target variable and keep the same configuration (CV of 3 folds and random search approach in the tuning process).
 
-```r
+``` r
 learner.wlog = lrn("regr.lightgbm",
@@
 )
-
 ```
 
-The best model using the logarithmic transformation had different hyperparameters relative to the previous experiment and obtained an RMSE of 34,293.33 on the test dataset.
-This result shows that the error of this model is greater than the best model without the transformation (31,745.54).
+The best model using the logarithmic transformation has different hyperparameters relative to the previous experiment and obtains an RMSE of 34,293.33 on the test data set.
+Thus, the error of this model is greater than the best model without the transformation (31,745.54).
 
@@
 
-Gradient Boosting also has a similar amount of hyperparameters as random forest and we decided to keep a similar configuration as the previous algorithm to find the best model with the hyperparameters (random search approach in the tuning process with 3-fold cross-validation).
-This first experiment with Gradient Boosting uses the target variable without logarithmic transformation.
+Gradient Boosting has a similar amount of hyperparameters as random forest and we use a similar configuration as the previous algorithm to find the best model with the hyperparameters (random search approach in the tuning process with 3-fold CV).
+The first model with Gradient Boosting uses the target variable without logarithmic transformation.
 
@@
 
-We tried many combinations of hyperparameters and found the best model that produced the least error.
-In Gradient Boosting we adjusted some hyperparameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
-Once the best model was found, it was evaluated on the test data set and an RMSE of 29,207.55 was obtained.
-The hyperparameters of the best model are shown in the following table.
+Our best model uses a combination of hyperparamters that produces the least error.
+We adjust some hyperparameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others.
+We evaluate the best model with the test data set and obtain an RMSE of 29,207.55.
+The hyperparameters of the best model are shown in figure 7.
 
-![Best hyperparameters for Gradient Boosting](includes/GB_hyper.png)
+```{r, echo=FALSE, out.width="50%", fig.cap="Best Hyperparameters for Gradient Boosting"}
+knitr::include_graphics("includes/GB_hyper.png")
+```
 
 Gradient Boosting also provides information on the importance of features during the modeling process in this experiment.
-The following figure shows the 20 most important variables according to the Gini metric.
+Figure 8 shows the 20 most important variables according to the Gini metric.
+`Gender` is after work status the most important variable, giving additional evidence for a gender pay gap.
 
@@
 
-Furthermore, we carried out an additional experiment with gradient boosting using the logarithm transformation on the target variable.
+Furthermore, we estimate a gradient boosting model using the logarithm transformation on the target variable.
 
-```r
+``` r
 learner.gb.wlog = lrn("regr.lightgbm",
@@
 
-The best model using the logarithmic transformation had a RMSE of 34,385.47 on the test dataset.
-This result shows that the error of this model is greater relative to the best model without the transformation (29,207.55).
+The best model using the logarithmic transformation has an RMSE of 34,385.47 on the test data set.
+Thus, the error of this model is greater relative to the best model without the transformation (29,207.55).
 These results show that in tree-based methods, Gradient Boosting has a dominant performance over Random Forest.
-In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used, but the errors of these models are larger than those of the linear models.
+In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used.
+Nevertheless, the errors of these models are larger than those of the linear models.
 
@@
 
-Generalized additive model (GAM) could be considered the next step of the linear model because it allows combining nonlinear functional forms of the features along with the linear part.
-This approach is based on relating the nonlinear predictors with a link function $g(.)$ of the expected values $E(y)$, and the following formulas offer more details about this approach.
+Generalized additive models (GAM) could be considered the next step of the linear model because they allow combining nonlinear functional forms of the features along with a linear part.
+This approach is based on relating the nonlinear predictors with a link function $g(.)$ of the expected values $E(y)$.
+The following formulas offer more details about this approach.
 
 $$
-
-y \sim \text{ExpoFam}(\mu, \text{etc})\\
-E(y) = \mu \\
+y \sim \text{ExpoFam}(\mu, \text{etc.})
+$$ $$
+E(y) = \mu
+$$ $$
 g(\mu) = b_0 + f(x_1) + f(x_2) + \dots + f(x_p)
-
-$$
-
-The main difference with Generalized Lineal Models (GLM) is that this approach allows linear features to aggregate smooth functions $f(.)$ of the required variables and thus establish nonlinear relationships between regressors and the dependent variable.
-This models uses the `gam` and `mlr3` packages to model for the modeling.
-The adjusted hyperparameter was `gamma` and is responsible for producing a smoother term, and a additional penalization on the smooth term was aggregated with the parameter `select`.
-In addition, we used a 3-fold cross-validation in the training stage for this algorithm.
-This experiment uses the logarithmic transformation of the target variable, the interaction variables, smooth terms for the features *age* and *prestg10* with cubic regression splines `cr` and a value of 9 for the knots `k` (number of sections into which a feature is divided), and the smooth tensor product for the variables *age*, *prestg10* and *child* uses a thin-plate regression spline `tp` and a knot value of 1.
+$$ The main difference with Generalized Lineal Models (GLM) is that this approach allows linear features to aggregate smooth functions $f(.)$ of the required variables and thus establish nonlinear relationships between regressors and the dependent variable.
+We use the `gam` and `mlr3` packages to model the GLM.
+The adjusted hyperparameter is `gamma` and is responsible for producing a smoother term.
+An additional penalization on the smooth term is aggregated with the parameter `select`.
+In addition, we use a 3-fold CV in the training stage for this algorithm.
+We use the logarithmic transformation of the target variable, the interaction variables, smooth terms for the features `age` and `prestg10` with cubic regression splines `cr` and a value of 9 for the knots `k` (number of sections into which a feature is divided).
+The smooth tensor product for the variables `age`, `prestg10` and `child` uses a thin-plate regression spline `tp` and a knot value of 1.
 The last smooth term is considered as the interaction between two features to predict the target variables.
 
-```r
-form.str.gam <- paste("log_realrinc ~ s(age, bs = 'cr', k = 9) + s(prestg10, bs = 'cr', k=9) + 
+``` r
+form.str.gam <- paste("log_realrinc ~ s(age, bs = 'cr', k = 9) 
+                                    + s(prestg10, bs = 'cr', k=9) + 
                         ti(age,prestg10) + ti(age,childs) +",
@@
 
-The best model obtained a RMSE of 25,856.76 on the test dataset.
-This result shows that the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82).
+The best model obtaines an RMSE of 25,856.76 on the test data set.
+Thus, the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82).
 In other words, it is 0.20% better than the initial linear model.
-The figure below exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.
+Figure 9 exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.
 
-![Approximate significance of smooth terms](includes/aprox_smooth.png)
+```{r, echo=FALSE, out.width="80%", fig.cap="Approximate Significance of Smooth Terms"}
+knitr::include_graphics("includes/aprox_smooth.png")
+```
 
@@
 
+```{=tex}
 \begin{tabular}{l c} 
@@
 \end{tabular}
+```
+\medskip
 
-\medskip
 Conclusively, the linear regression yields the lowest RMSE and thus appears to be the best prediction method.
 Still, it delivers negative predictions of the wage.
-Therefore, we prefer the logistic regression as the best prediction method.
+Therefore, we prefer the logistic regression.
+Comparing the RMSE of the logistic model to the one of the GAM, the GAM (which presents a combination of linear and nonlinear functional forms of the features) produces the lowest RMSE.
+Hence, the GAM presents the best prediction method for real wages (in our data set).
+
+Moreover, the different prediction methods consider `gender` as an important determinant for predicting the individuals' real wage.
+Still, as we do not use any causal methods other unobserved factors might influence the wage more than `gender`.
+
+\pagebreak
+
+\addcontentsline{toc}{section}{References}
+\printbibliography[title = References]
+\cleardoublepage
+
+\begin{refsection}
+\nocite{R-base}
+\nocite{R-Studio}
+
+\end{refsection}

### 2023-08-31 Paper/paper.tex
proofread papper

@@
 % use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 % use microtype if available
@@
 %% citation setup
-\usepackage{csquotes}
+%\usepackage{csquotes}
 
@@
   \postdate{\par}
-  \date{today}
+  \date{31.08.2023}
 
@@
   \begin{minipage}[t]{0.7\textwidth}
-  \hspace{1cm}tomorrow
+  \hspace{1cm}31.08.2023
   \end{minipage}
@@
 
-  123456
+  3104233
 
@@
 
-PhD Economics
+Ph.D. Economics
 
@@
 
-  Summer Term 2023
+  Summer Term 2028
 
@@
 
-  John Doe
+  Alex Amaguaya
 
-  234567
+  242833
 
-  john.doe@web.de
+  alex.amaguaya@tu-dortmund.de
 
-  M.Sc. Economics
+  M.Sc. Econometrics
 
-  \nth{4}
+  \nth{2}
 
-  Summer Term 2020
+  Summer Term 2023
 
@@
 The prediction of individuals' wages has been an issue in previous
-literature. As wages are relevant to companies they want to know how
+literature. As wages are relevant to companies they want to know how to
 predict wages \autocite{Chakraborti}. Some papers make use of
@@
 showed better performance compared to a traditional algorithm such as
-SVM. Furthermore, \autocite{BonaccoltoToepfer1612} suggests the use of a
-penalization approach for high dimensional problems in his research
-about Gender wage inequality.
+support vector machines (SVM). Furthermore,
+\textcite{BonaccoltoToepfer1612} suggest the use of a penalization
+approach for high dimensional problems in his research about gender wage
+inequality.
 
@@
 \item
+  Random Forest
+\item
   Gradient Boosting
 \item
-  Random Forest
+  Generalized Additive Model
 \end{itemize}
@@
 discarded to avoid generating a bias in the modeling and therefore we
-decided to eliminate these records from the original data set. After
-this, the dataset contains 33,702 observations. Furthermore, it is
-impossible to deduce the occupation for some observations
-(\texttt{occ10\ =\ 9997}) and we remove these from the data set. At the
-end, the final data set has 33,244 observations, and we use it for the
-modeling process.
+eliminate these records from the original data set. After this, the data
+set contains 33,702 observations. Furthermore, it is impossible to
+deduce the occupation for some observations (\texttt{occ10\ =\ 9997})
+and we remove these from the data set. Finally, the data set has 33,244
+observations and we use it for the modeling process.
 
-This data set still had some missing values for the regressors, e.g.,
-number of children, age, marital status, degree level, and others. We
-use the \texttt{mice} package with five variables for the imputation.
-The features \emph{age, occupational prestige and number of children}
-are imputed using the pmm (\texttt{predictive\ mean\ matching}) method.
-We use the polytomous logistic regression method for the factor
-variables (\emph{education and marital status}). The imputation process
-generates 6 different data sets. We apply the estimation of the mean and
-the majority vote to aggregate the data sets into a single data set. The
-mean estimation is applied for the numeric variables and majority vote
-for the categorical variables..
+This data set still has some missing values for the regressors, e.g.,
+number of children, age, marital status, degree level, and others. Thus,
+we use the \texttt{mice} package with five variables for the imputation
+process. The features \emph{age, occupational prestige and number of
+children} are imputed using the predictive mean matching (pmm) method,
+and we use the polytomous logistic regression method for the factor
+variables (\emph{education and marital status}). As we use a value of 6
+in the \texttt{m} parameter of the \texttt{mice} function, the
+imputation process generates 6 different data sets. This parameter
+refers to the number of imputed data sets. After this, we apply the
+calculation of the mean and the majority vote to aggregate the data sets
+into one. Mean estimation is applied for numerical variables and
+majority voting for categorical variables
 
@@
 
-After the imputation process, we create the interaction variables.
-First, we estimate some interaction features using only the numerical
-variables. Then, we transform the numerical variables into categorical
-variables by dividing them into different ranges (e.g., age between 18
-and 30, age between 31 and 50, etc.). We transform these new categorical
-variables and the initial categorical variables (\emph{marital status,
-education, etc.}) into dummy variables. And finally, the interaction
-variables are estimated with the group of dummy variables mentioned
-above. The inclusion of the interaction of the numeric variables are
-validate using a linear model, and the results show that the p- values
-of the parameters for the interactions are less than 0.05.
+After the imputation process, we create interaction variables. First, we
+estimate some interaction features using only the numerical variables.
+Then, we transform the numerical variables
+(\texttt{age,\ childs\ and\ prestg10}) into categorical variables by
+separating them into ranges. The ranges created of these variables were:
+\texttt{age\ between\ 18\ and\ 30,\ age\ between\ 31\ and\ 50},
+\texttt{age\ greater\ than\ 50;\ childs\ between\ 0\ and\ 2,\ childs\ between\ 3\ and\ 5,\ childs\ more\ than\ 5;\ prestg10\ between\ 16\ and\ 30,\ prestg10\ between\ 31\ and\ 50,\ prestg10\ greater\ than\ 50}.
+After this, we transform these new categorical variables and the initial
+categorical variables (\texttt{marital\ status}, \texttt{education},
+etc.) into dummy variables. Finally, we estimate the interaction
+variables with the group of dummy variables mentioned above (the
+interaction consists of the multiplication between two dummy variables).
+The inclusion of the interactions of the numeric variables is validate
+using a linear model, and the results show that the p-values of the
+parameters for the interactions are below 0.05.
 
@@
 correlation coefficient with some numeric variables in order to support
-the previous results. The correlation results are shown in the following
-figure.
+the previous results. The correlation results are exposed in figure 2,
+and show that \texttt{prestg10}, \texttt{age} and
+\texttt{interaction\ between\ age\ and\ prest10} have the highest
+positive correlation with \texttt{realrinc}, meanwhile the dummy
+variable \texttt{female} has a moderate negative correlation with the
+target variable.
 
@@
 
-After the data preparation and feature engineering processes, we
-construct a final data set with 33,244 observations and 421 features.
-Due to the large number of interaction variables, we decided to reduce
-them and use only the interaction between a group of variables
-(\emph{occrecode, educcat and prestg10}). After reduction, we develop
-the modeling process with 100 variables (numerical and interaction
-variables). The following figure shows the overview of the group of
-variables that were estimated and used in the models.
+As the correlation assesses the linear relationship between two
+variables and it only uses numerical features, we additionally estimate
+the Predictive Power Score (PPS). It overcomes these limitations and
+allows us to assess the predictive power of a regressor on a target
+variable. This measure is a normalized metric and its values are between
+0 (no predictive power) and 1 (perfect predictive power). The PPS is
+calculated by comparing the performance of a model that uses a regressor
+\texttt{x} to predict target variable \texttt{y} versus the performance
+of a base model (naive). When the target variable is a numerical feature
+it is common to use a Decision Tree Regressor to predict \texttt{y}
+based on \texttt{x}, while in categorical cases a Decision Tree
+Classifier is used. On the other hand, base models are often defined as
+mean prediction or random or modal prediction for regression and
+classification problems, respectively. The PPS results are presented in
+figure 3, and show that \texttt{prestg10}, \texttt{wrkstat},
+\texttt{educcat} and \texttt{occrecode} have the highest predictive
+power of \texttt{realrinc} relative to the other features. These results
+provide some insights in the data exploration process, but we must
+beware of the limitations of this approach.
+
+\begin{figure}
+\centering
+\includegraphics{includes/pps.png}
+\caption{PPS Results}
+\end{figure}
+
+After the data preparation and feature engineering processes, we have a
+final data set with 33,244 observations and 421 features. Due to the
+large number of interaction variables, we reduce them and use only the
+interaction between a group of variables (\texttt{occrecode},
+\texttt{educcat} and categorical variable of \texttt{prestg10}). After
+this, we develop the modeling process with 100 variables (numerical and
+interaction variables). Figure 4 shows the overview of the group of
+estimated variables that are used in the models.
 
@@
 
-As the dataset contains many variables (361) and even our selection of
+As the data set contains many variables (361) and even our selection of
 dummy variables for the right-hand side of the regression leaves us with
@@
 multiplied with a tuning parameter \(\lambda\). We estimate \(\lambda\)
-using cross-validation.
+using CV.
 
@@
 According to the literature review, tree-based models perform
-considerably well in range in predicting income ranges (classification
-problems). For this reason, we use the random forest algorithm for
-modeling. This experiment use the income without logarithm
-transformation, the initial features and some interaction variables.
+considerably well in predicting income ranges (classification problems)
+\autocite{Chakrabarty2018}. For this reason, we use the random forest
+algorithm for modeling. For this method we use income without logarithm
+transformation, initial variables and some interaction variables. Due to
+the amount of hyperparameters in random forest, we use the
+\texttt{lightgbm} and \texttt{mlr3} packages to model and tune the
+hyperparameter in a time-optimal way. We tune the hyperparameters
+\texttt{max\_depth}, \texttt{num\_leaves}, \texttt{num\_iterations},
+etc. to find the best hyperparameters that produce the lowest error. The
+search space of the hyperparameters is showed below. In addition, we use
+a random search approach in the tuning process with a CV of 3 folds.
 
@@
   \AttributeTok{measures =}\NormalTok{ msr\_ce,}
-  \AttributeTok{term\_evals =} \DecValTok{150}\NormalTok{,}
+  \AttributeTok{term\_evals =} \DecValTok{300}\NormalTok{,}
   \AttributeTok{store\_models =} \ConstantTok{FALSE}
@@
 
-Due to the number of hyper parameters in random forest, we use the
-\texttt{lightgbm} and \texttt{mlr3} packages to model and tune the hyper
-parameter in a time-optimal way. After many attempts, we found the best
-hyper parameters that produced the lowest error. The hyper parameters
-tuned were \texttt{max\_depth}, \texttt{num\_leaves},
-\texttt{num\_iterations} and others. In addition, we used a random
-search approach in the tuning process with a cross-validation of 3. Once
-we found the model along with the hyper parameters with lowest error in
-the cross-validation, we proceeded to evaluate this model on the test
-data set and obtained a RMSE of 31,745.54. The hyper parameters of the
-best model are shown in the following table.
+After identifying the model along with the hyperparameters with lowest
+error in CV, we proceed to evaluate this model on the test data set and
+obtain an RMSE of 31,745.54. The hyperparameters of the best model are
+shown in figure 5.
 
-{[}best hyper parameter of RF{]}
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{includes/RF_hyper} \caption{Best Hyperparameters for Random Forest}\label{fig:unnamed-chunk-1}
+\end{figure}
 
-In addition, we can analyze the importance of the features within the
-modeling process using random forest, and these results are shown in the
-following table.
+In addition, we analyze the features importance within the modeling
+process using random forest. These results are exposed in the figure 6.
 
-{[}feature importance of RF{]}
+\begin{figure}
+\centering
+\includegraphics{includes/RF_importance.png}
+\caption{Top 20 of Features Importance for Random Forest}
+\end{figure}
+
+Unlike the linear models presented previously, random forest does not
+predict negative values when logarithmic transformation is not used.
+Still, we additionally use a random forest with the logarithmic
+transformation in the target variable and keep the same configuration
+(CV of 3 folds and random search approach in the tuning process).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{learner.wlog }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
+              \AttributeTok{boosting =} \StringTok{"rf"}\NormalTok{,}
+              \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
+              \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{7}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{120}\NormalTok{,}\DecValTok{140}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{max\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{45}\NormalTok{,}\DecValTok{60}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{50}\NormalTok{,}\DecValTok{90}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{20}\NormalTok{,}\DecValTok{35}\NormalTok{,}\DecValTok{1}\NormalTok{)), }
+              \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{35}\NormalTok{,}\DecValTok{50}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{,}\FloatTok{0.7}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{bagging\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.6}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{bagging\_freq =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+              \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.8}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+              \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
+              \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
+              \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
+              \AttributeTok{num\_threads =} \DecValTok{5}\NormalTok{,}
+              \AttributeTok{seed =} \DecValTok{123}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The best model using the logarithmic transformation has different
+hyperparameters relative to the previous experiment and obtains an RMSE
+of 34,293.33 on the test data set. Thus, the error of this model is
+greater than the best model without the transformation (31,745.54).
 
@@
 
-Gradient Boosting also has many hyper parameters like random forest and
-we decided to apply a similar approach to the previous algorithm to find
-the best model with the hyper parameters (random search approach in the
-tuning process with a cross-validation of 3). Furthermore, the
-experiments with Gradient Boosting use the target variable without
-logarithm transformation.
+Gradient Boosting has a similar amount of hyperparameters as random
+forest and we use a similar configuration as the previous algorithm to
+find the best model with the hyperparameters (random search approach in
+the tuning process with 3-fold CV). The first model with Gradient
+Boosting uses the target variable without logarithmic transformation.
 
@@
                  \AttributeTok{learning\_rate =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.03}\NormalTok{, }\FloatTok{0.04}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
-                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)), }\CommentTok{\#40}
+                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                  \AttributeTok{lambda\_l1 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
@@
   \AttributeTok{measures =}\NormalTok{ msr\_ce,}
-  \AttributeTok{term\_evals =} \DecValTok{170}\NormalTok{,}
+  \AttributeTok{term\_evals =} \DecValTok{300}\NormalTok{,}
   \AttributeTok{store\_models =} \ConstantTok{FALSE}
@@
 
-We tried many experiments and found the best hyper parameters that
-produced the lowest error. In Gradient Boosting modeling we adjusted the
-hyper parameters such as \texttt{max\_depth}, \texttt{num\_leaves},
-\texttt{min\_data\_in\_leaf}, \texttt{feature\_fraction} and others.
-Once the best model was found, it obtained an RMSE of 29,207.55 on the
-test dataset. The hyper parameters of the best model are shown in the
-following table.
+Our best model uses a combination of hyperparamters that produces the
+least error. We adjust some hyperparameters such as \texttt{max\_depth},
+\texttt{num\_leaves}, \texttt{min\_data\_in\_leaf},
+\texttt{feature\_fraction} and others. We evaluate the best model with
+the test data set and obtain an RMSE of 29,207.55. The hyperparameters
+of the best model are shown in figure 7.
 
-{[}best hyper parameter of GB{]}
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{includes/GB_hyper} \caption{Best Hyperparameters for Gradient Boosting}\label{fig:unnamed-chunk-2}
+\end{figure}
 
-Furthermore, the following table shows the importance of the features
-within the modeling process using Gradient Boosting.
+Gradient Boosting also provides information on the importance of
+features during the modeling process in this experiment. Figure 8 shows
+the 20 most important variables according to the Gini metric.
+\texttt{Gender} is after work status the most important variable, giving
+additional evidence for a gender pay gap.
 
-{[}feature importance of GB{]}
+\begin{figure}
+\centering
+\includegraphics{includes/GB_importance.png}
+\caption{Top 20 of Features Importance for Gradient Boosting}
+\end{figure}
 
-These results show that in tree-based methods, the Gradient Boosting
-performance overcomes Random Forest, in other words, the error of
-Gradient Boosting is approximately 8\% lower than the error of Random
-Forest. In addition, the errors of these models are larger than the
-linear models.
+Furthermore, we estimate a gradient boosting model using the logarithm
+transformation on the target variable.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{learner.gb.wlog }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
+                 \AttributeTok{boosting =} \StringTok{"gbdt"}\NormalTok{,}
+                 \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
+                 \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{12}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{max\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{35}\NormalTok{,}\DecValTok{60}\NormalTok{,}\DecValTok{1}\NormalTok{)), }
+                 \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{19}\NormalTok{,}\DecValTok{30}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{30}\NormalTok{,}\DecValTok{50}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{20}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
+                 \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.8}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+                 \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.7}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
+                 \AttributeTok{learning\_rate =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\FloatTok{0.01}\NormalTok{)), }
+                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{60}\NormalTok{,}\DecValTok{80}\NormalTok{,}\DecValTok{1}\NormalTok{)), }
+                 \AttributeTok{lambda\_l1 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
+                 \AttributeTok{lambda\_l2 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{, }\FloatTok{0.45}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
+                 \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
+                 \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
+                 \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
+                 \AttributeTok{num\_threads =} \DecValTok{5}\NormalTok{,}
+                 \AttributeTok{seed =} \DecValTok{123}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The best model using the logarithmic transformation has an RMSE of
+34,385.47 on the test data set. Thus, the error of this model is greater
+relative to the best model without the transformation (29,207.55). These
+results show that in tree-based methods, Gradient Boosting has a
+dominant performance over Random Forest. In other words, the error of
+Gradient Boosting is about 8\% smaller than the error of Random Forest
+when the logarithmic transformation is not used. Nevertheless, the
+errors of these models are larger than those of the linear models.
+
+\hypertarget{generalized-additive-model}{%
+\subsection{Generalized Additive
+Model}\label{generalized-additive-model}}
+
+Generalized additive models (GAM) could be considered the next step of
+the linear model because they allow combining nonlinear functional forms
+of the features along with a linear part. This approach is based on
+relating the nonlinear predictors with a link function \(g(.)\) of the
+expected values \(E(y)\). The following formulas offer more details
+about this approach.
+
+\[
+y \sim \text{ExpoFam}(\mu, \text{etc.})
+\] \[
+E(y) = \mu
+\] \[
+g(\mu) = b_0 + f(x_1) + f(x_2) + \dots + f(x_p)
+\] The main difference with Generalized Lineal Models (GLM) is that this
+approach allows linear features to aggregate smooth functions \(f(.)\)
+of the required variables and thus establish nonlinear relationships
+between regressors and the dependent variable. We use the \texttt{gam}
+and \texttt{mlr3} packages to model the GLM. The adjusted hyperparameter
+is \texttt{gamma} and is responsible for producing a smoother term. An
+additional penalization on the smooth term is aggregated with the
+parameter \texttt{select}. In addition, we use a 3-fold CV in the
+training stage for this algorithm. We use the logarithmic transformation
+of the target variable, the interaction variables, smooth terms for the
+features \texttt{age} and \texttt{prestg10} with cubic regression
+splines \texttt{cr} and a value of 9 for the knots \texttt{k} (number of
+sections into which a feature is divided). The smooth tensor product for
+the variables \texttt{age}, \texttt{prestg10} and \texttt{child} uses a
+thin-plate regression spline \texttt{tp} and a knot value of 1. The last
+smooth term is considered as the interaction between two features to
+predict the target variables.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{form.str.gam }\OtherTok{\textless{}{-}} \FunctionTok{paste}\NormalTok{(}\StringTok{"log\_realrinc \textasciitilde{} s(age, bs = \textquotesingle{}cr\textquotesingle{}, k = 9) }
+\StringTok{                                    + s(prestg10, bs = \textquotesingle{}cr\textquotesingle{}, k=9) + }
+\StringTok{                        ti(age,prestg10) + ti(age,childs) +"}\NormalTok{,}
+                       \FunctionTok{paste}\NormalTok{( x\_cols\_dummys, }\AttributeTok{collapse =} \StringTok{" + "}\NormalTok{) )}
+\NormalTok{form.format.gam }\OtherTok{\textless{}{-}} \FunctionTok{as.formula}\NormalTok{(form.str.gam)}
+\NormalTok{learner.gam.wlog.smooth.tensor }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.gam"}\NormalTok{,}
+                                     \AttributeTok{family =} \StringTok{\textquotesingle{}gaussian\textquotesingle{}}\NormalTok{,}
+                                     \AttributeTok{select =} \ConstantTok{TRUE}\NormalTok{,}
+                                     \AttributeTok{gamma =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{3}\NormalTok{,}\FloatTok{0.5}\NormalTok{)),}
+                                     \AttributeTok{formula =}\NormalTok{ form.format.gam}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The best model obtaines an RMSE of 25,856.76 on the test data set. Thus,
+the error of this model is slightly lower in relation to the linear
+model that also uses the logarithmic transformation (25,909.82). In
+other words, it is 0.20\% better than the initial linear model. Figure 9
+exposes the approximate significance of the smooth terms, and the
+results show that all the smooth terms used in the model have a p-value
+of less than 0.01.
+
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{includes/aprox_smooth} \caption{Approximate Significance of Smooth Terms}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 appears to be the best prediction method. Still, it delivers negative
-predictions of the wage. Therefore, we prefer the logistic regression as
-the best prediction method.
+predictions of the wage. Therefore, we prefer the logistic regression.
+Comparing the RMSE of the logistic model to the one of the GAM, the GAM
+(which presents a combination of linear and nonlinear functional forms
+of the features) produces the lowest RMSE. Hence, the GAM presents the
+best prediction method for real wages (in our data set).
+
+Moreover, the different prediction methods consider \texttt{gender} as
+an important determinant for predicting the individuals' real wage.
+Still, as we do not use any causal methods other unobserved factors
+might influence the wage more than \texttt{gender}.
+
+\pagebreak
+
+\addcontentsline{toc}{section}{References}
+\printbibliography[title = References]
+\cleardoublepage
+
+\begin{refsection}
+\nocite{R-base}
+\nocite{R-Studio}
+
+\end{refsection}
 

### 2023-08-31 code/4_LassoRidgeElasticNet.R

@@ -83,7 +83,6 @@ lasso <- train(formula, data = train,
                 tuneGrid = expand.grid(alpha = 1, lambda = seq(0,1,0.001)))
 caret::RMSE(pred = exp(predict(lasso, validation)), obs = validation$realrinc) #25918.25
 
-
 #Ridge
 set.seed(123)
 ridge <- train(formula, data = train,

### 2023-08-31 Paper/paper.tex
minor improvements

@@
 
-In this paper we predict individuals' wages using data from the General
-Social Survey (GSS, 1974-2018). The variable of interest is
+In this paper we predict individuals' incomes using data from the
+General Social Survey (GSS, 1974-2018). The variable of interest is
 \texttt{realrinc} which varies between 227\$ and
-480,145\$\footnote{In constant 1986 USD.}. Hence, we can use regression
-techniques to predict individuals' wages. Additionally, we can consider
-whether \texttt{realrinc} varies by gender and if so, how much.
+480,145\$\footnote{In constant 1986 USD.}. Hence, we can employ
+regression techniques to predict individuals' incomes. Additionally, we
+can consider whether \texttt{realrinc} varies by gender and if so, how
+much.
 
@@
 
-The prediction of individuals' wages has been an issue in previous
+The prediction of individuals' incomes has been an issue in previous
 literature. As wages are relevant to companies they want to know how to
 predict wages \autocite{Chakraborti}. Some papers make use of
-classification techniques to predict wages \autocite{Chakraborti}. As we
-have numeric values of real wages and not a categorical variable we do
-not make use of classification techniques but prefer linear regression
-techniques such as OLS and Lasso regressions as well as random forests.
-We prefer these techniques as they handle categorical variables and
-outliers well, which are features of our data \autocite{Cutler}. Wages
-were already predicted using random forests by \autocite{Eichinger}.
-They find that using random forests results in better predictions than
-using linear models. Additionally, they show that gender barely effects
-the wage predictions. Likewise, the boosting methods used by
-\autocite{Chakrabarty2018} in salary level classification problems
-showed better performance compared to a traditional algorithm such as
-support vector machines (SVM). Furthermore,
-\textcite{BonaccoltoToepfer1612} suggest the use of a penalization
-approach for high dimensional problems in his research about gender wage
-inequality.
+classification techniques to predict wages \autocite{Chakraborti}. Since
+we have numeric values representing real wages and not a categorical
+variable we do not make use of classification techniques but prefer
+linear regression techniques such as OLS and Lasso regressions as well
+as random forests. We prefer these techniques as they handle categorical
+variables and outliers well, which are features of our data
+\autocite{Cutler}. Wages were already predicted using random forests by
+\autocite{Eichinger}. They find that using random forests results in
+better predictions than using linear models. Additionally, they show
+that gender barely effects the wage predictions. Likewise, the boosting
+methods used by \autocite{Chakrabarty2018} in salary level
+classification problems showed better performance compared to a
+traditional algorithm such as support vector machines (SVM).
+Furthermore, \textcite{BonaccoltoToepfer1612} suggest the use of a
+penalization approach for high dimensional problems in his research
+about gender wage inequality.
 
@@
 in the \texttt{m} parameter of the \texttt{mice} function, the
-imputation process generates 6 different data sets. This parameter
+imputation process generates six different data sets. This parameter
 refers to the number of imputed data sets. After this, we apply the
@@
 Classifier is used. On the other hand, base models are often defined as
-mean prediction or random or modal prediction for regression and
-classification problems, respectively. The PPS results are presented in
-figure 3, and show that \texttt{prestg10}, \texttt{wrkstat},
-\texttt{educcat} and \texttt{occrecode} have the highest predictive
-power of \texttt{realrinc} relative to the other features. These results
-provide some insights in the data exploration process, but we must
-beware of the limitations of this approach.
+mean, random or modal prediction for regression and classification
+problems, respectively. The PPS results are presented in figure 3, and
+show that \texttt{prestg10}, \texttt{wrkstat}, \texttt{educcat} and
+\texttt{occrecode} have the highest predictive power of
+\texttt{realrinc} relative to the other features. These results provide
+some insights in the data exploration process, but we must beware of the
+limitations of this approach.
 
@@
 large number of interaction variables, we reduce them and use only the
-interaction between a group of variables (\texttt{occrecode},
+interactions among a selected group of variables (\texttt{occrecode},
 \texttt{educcat} and categorical variable of \texttt{prestg10}). After
@@
 results show that in tree-based methods, Gradient Boosting has a
-dominant performance over Random Forest. In other words, the error of
+superior performance over Random Forest. In other words, the error of
 Gradient Boosting is about 8\% smaller than the error of Random Forest
@@
 model that also uses the logarithmic transformation (25,909.82). In
-other words, it is 0.20\% better than the initial linear model. Figure 9
-exposes the approximate significance of the smooth terms, and the
-results show that all the smooth terms used in the model have a p-value
-of less than 0.01.
+other words, it performs 0.20\% better than the initial linear model.
+Figure 9 exposes the approximate significance of the smooth terms, and
+the results show that all the smooth terms used in the model have a
+p-value of less than 0.01.
 

### 2023-08-31 code/0_code.R
common code script

@@ -1,0 +1,781 @@
+# set wd
+setwd("/home/user/Downloads/info_tud/statistical_learning_UDE/gender_pay_gap_project")
+# import functions file
+source('code/utils_functions.R')
+setwd("C:/Users/lbergmann/OneDrive - RWI–Leibniz-Institut für Wirtschaftsforschung e.V/Dokumente/Promotion/StatisticalLearning/gender_pay_gap_project_1")
+
+# packages
+library(dplyr)
+library(GGally)
+library(mice)
+library(beepr)
+library(collapse)
+library(parallel)
+library(caret)
+library(ggplot2)
+library(stringr)
+library(ISLR)
+library(tidyverse)
+library(lightgbm)
+library(Matrix)
+library(parallel)
+library(sgd)
+library(mlr3)
+library(mlr3learners)
+library(mlr3tuning)
+library(data.table)
+library(stargazer)
+# remotes::install_github("mlr-org/mlr3extralearners@*release")
+library(mlr3extralearners)
+
+
+# import ds
+ds <- read.csv('data/gss_wages_train.csv')
+summary(ds)
+ds1 <- ds
+cat("the dataset has ", nrow(ds), 'rows, and ', round(sum(is.na(ds$realrinc))/nrow(ds)*100,2), "% of missing values for the outcome \n")
+
+#create factor variables
+ds$educcat <- as.factor(ds$educcat)
+ds$maritalcat <- as.factor(ds$maritalcat)
+ds$wrkstat <- as.factor(ds$wrkstat)
+ds$gender <- as.factor(ds$gender)
+ds$occrecode <- as.factor(ds$occrecode)
+
+# filtered dataset, not using obs, with missing value for outcome
+ds_filter  <- ds[!is.na(ds$realrinc),] #filtered dataset
+summary(ds_filter)
+
+cat("the dataset has ", nrow(ds_filter), 'rows \n')
+summary(ds_filter)
+
+# NA analysis for filtered dataset: 
+# 33066 obs for which we have all variables, 
+# 382 obs miss prestige
+# 82 miss occupation and prestige
+# 71 miss children
+md.pattern(ds_filter)
+
+# NA analysis: children
+ds_child <- ds_filter %>% filter(is.na(childs))
+summary(ds_child)
+# those for whom we don't know whether they have children are mainly in their 30s
+# male, studied till the high school, Never Married, and work full-time
+
+# NA analysis: prestige
+ds_prest <- ds_filter %>% filter(is.na(prestg10))
+summary(ds_prest)
[…]

### 2023-09-22 Paper/paper.Rmd

@@
 
-In this paper we predict individuals' wages using data from the General Social Survey (GSS, 1974-2018).
+In this paper we predict individuals' incomes using data from the General Social Survey (GSS, 1974-2018).
 The variable of interest is `realrinc` which varies between 227\$ and 480,145\$\footnote{In constant 1986 USD.}.
-Hence, we can use regression techniques to predict individuals' wages.
+Hence, we can employ regression techniques to predict individuals' incomes.
 Additionally, we can consider whether `realrinc` varies by gender and if so, how much.
@@
 
-The prediction of individuals' wages has been an issue in previous literature.
+The prediction of individuals' incomes has been an issue in previous literature.
 As wages are relevant to companies they want to know how to predict wages [@Chakraborti].
 Some papers make use of classification techniques to predict wages [@Chakraborti].
-As we have numeric values of real wages and not a categorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests.
+Since we have numeric values representing real wages and not a categorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests.
 We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler].
@@
 The features *age, occupational prestige and number of children* are imputed using the predictive mean matching (pmm) method, and we use the polytomous logistic regression method for the factor variables (*education and marital status*).
-As we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets.
+As we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates six different data sets.
 This parameter refers to the number of imputed data sets.
@@
 When the target variable is a numerical feature it is common to use a Decision Tree Regressor to predict `y` based on `x`, while in categorical cases a Decision Tree Classifier is used.
-On the other hand, base models are often defined as mean prediction or random or modal prediction for regression and classification problems, respectively.
+On the other hand, base models are often defined as mean, random or modal prediction for regression and classification problems, respectively.
 The PPS results are presented in figure 3, and show that `prestg10`, `wrkstat`, `educcat` and `occrecode` have the highest predictive power of `realrinc` relative to the other features.
@@
 After the data preparation and feature engineering processes, we have a final data set with 33,244 observations and 421 features.
-Due to the large number of interaction variables, we reduce them and use only the interaction between a group of variables (`occrecode`, `educcat` and categorical variable of `prestg10`).
+Due to the large number of interaction variables, we reduce them and use only the interactions among a selected group of variables (`occrecode`, `educcat` and categorical variable of `prestg10`).
 After this, we develop the modeling process with 100 variables (numerical and interaction variables).
@@
 Thus, the error of this model is greater relative to the best model without the transformation (29,207.55).
-These results show that in tree-based methods, Gradient Boosting has a dominant performance over Random Forest.
+These results show that in tree-based methods, Gradient Boosting has a superior performance over Random Forest.
 In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used.
@@
 Thus, the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82).
-In other words, it is 0.20% better than the initial linear model.
+In other words, it performs 0.20% better than the initial linear model.
 Figure 9 exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.
